---
title: "RAP Skeleton"
lang: "en"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_download: no
    css: "style.css"
    includes:
      in_header:
        - "meta.html"
        - "consent_head.html"
      after_body:
        - "cookie_banner.html"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../outputs/",
      output_file = paste0(
        xfun::sans_ext(input),
        "_",
        format(
          Sys.time(),
          "%d-%m-%Y_%H%M"
        ),
        ".html"
      ),
      envir = globalenv()
    )
  })
params:
  annotations: true
---

<div id="cookie-banner"></div>

```{r settings, include=FALSE}
# this chunk sets chunk options for all chunks in this file
knitr::opts_chunk$set(
  message = FALSE,
  echo = FALSE,
  warning = FALSE
)
```

```{r setup, warnings = FALSE}
library(here)

source(
  here(
    "code",
    "data_prep.R"
  )
)
```

<!-- message to rotate screen when width is too low -->

<div data-nics-theme="`r nics_theme`">
<!-- define header landmark area for accessibility -->
<header>
<!-- calling the banner code -->
`r f_banner(title = "NISRA Technology and Support Lab RAP Skeleton Demo Report")`

<!-- Header section -->
`r f_header()`
</header>

<!-- Introduction section -->

<div id="main" role="main">
`r f_borderline()`  <!-- calling the borderline code -->
 
<!-- Single Rmd -vs- 'sub-reports' -->
<!-- Depending on your requirements, you can edit this document - add text,
charts, links etc. directly into report.Rmd and knit -->
<!-- For larger projects with multiple large sections, or when you need many
contributors to edit it simultaneously, a 'sub-report' approach is
available -->
<!-- With 'sub-reports', you can dynamically reference other .Rmd content into
this file, similar to a cell reference in Excel -->
<!-- The demo code (code/demo/demo_report.Rmd) makes use of this (Note: you
can have a hybrid e.g. Key points is hard-coded and sub-reports for other
sections  -->

<!-- Example: code/demo/demo_report.Rmd: -->
<!-- ```{r introduction_sub_report, child='02_introduction.Rmd' }       -->
<!-- ```                                                                -->
<!-- 'introduction_sub_report' is the name of the section               -->
<!-- child='02_introduction.Rmd' is a reference to the file you want
to include  -->


## 1. Section header (h2)

Section text.

### Subsection heading (h3)

Subsection text.    

## Contact Details

`r f_contact()`  

### Accessibility contact (h3)

Please contact Dissemination Branch for assistance with accessibility
requirements or alternative formats. Contact details are:

**Email:** info@nisra.gov.uk

**Telephone:** +44 (0)300 200 7836

Dissemination Branch  
NISRA  
Colby House  
Stranmillis Court  
BELFAST  
BT9 5RR  
  
  

<!-- closing main div -->
</div>
<!-- footer landmark area is defined in footer function -->


```{r render-footer, results='asis', echo=FALSE}
f_footer()
```

<!-- message to rotate screen if it is below 400px width -->
<div id="rotate-screen">
  <p style="color: white;">Please rotate your screen for optimal experience</p>
</div>

<script>
  // Create array of alt texts for map shapes
  const map_alt = `r if (exists("map_alt")) {map_alt} else {"[]"}`;
</script>

`r if (params$annotations) '<script id="annotation-script" src="annotation.js"></script>'`

```{js}


// Accessibility function to remove map shapes from tab order, add tabs to table
// of contents and add empty alt text to map legend

function tabOrder() {

  // Removes tabs from individual map shapes and appends alt texts

  var shapes = document.getElementsByClassName(
    'leaflet-marker-icon leaflet-zoom-animated leaflet-interactive'
  );

  for (var i = 0; i < shapes.length; i++) {
    shapes[i].tabIndex = -1;
    shapes[i].alt = map_alt[i];
  }

  // Make side menu tabbable

  let headings = document.getElementsByClassName(
    'tocify-item list-group-item'
  );

  let sections = document.getElementsByClassName('section');

  for (let i = 0; i < headings.length; i++) {
    headings[i].innerHTML =
      "<a href='#" +
      sections[i].id +
      "'>" +
      headings[i].innerHTML +
      "</a>";
  }

  // Remove white space at bottom

  const space_observer = new MutationObserver(() => {
    let space = document.getElementsByClassName("tocify-extend-page");

    for (let i = 0; i < space.length; i++) {
      space[i].remove();
    }
  });

  space_observer.observe(
    document.getElementsByClassName("main-container")[0],
    {
      childList: true,
      subtree: true
    }
  );

  // Hide entire legend from screen readers

  var legendContainer = document.querySelector('.info.legend');

  if (legendContainer) {
    legendContainer.setAttribute('aria-hidden', 'true');
  }

  // Add empty alt text to each legend image

  var legendImages = document.querySelectorAll('.info.legend img');

  for (var j = 0; j < legendImages.length; j++) {
    legendImages[j].setAttribute('alt', '');
    legendImages[j].setAttribute(
      'aria-hidden',
      'true'
    ); // Extra assurance
  }
}

// Instruct web browser to run above function when loaded
window.onload = tabOrder;


```

<!-- close data-nics-theme div-->
</div>
