---
title: "04 Tables Sub Report"
output: html_document
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_dir = "../demo/demo_outputs/",
    output_file = paste0(
      xfun::sans_ext(input),
      "_",
      format(Sys.time(), "%d-%m-%Y_%H%M"),
      ".html"
    ),
    envir = globalenv()
  )
})
---

```{r tables_settings, include=FALSE}
# Set chunk options for all chunks in this file.

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

### Good practice
The Government Analysis function makes a distinction between demonstration
tables and reference tables. Demonstration tables are used to reinforce a point
made in the text of a publication. They are smaller and less detailed than
reference tables.

#### Accessible tables

Keep the structure of tables simple so everyone can use them. This is especially
important for people who use screen readers.

To ensure tables are clear and easy to read:

* do not merge cells
* do not leave rows or columns blank
* give every column a heading
* make sure all headings and labels are easy to see
* round numbers in the same way throughout the table
* align text to the left and numbers to the right
* use commas in large numbers, for example 1,000
* avoid abbreviations or explain them the first time you use them
* add labels such as %, £, or millions to the heading instead of repeating them
in each cell
* do not use symbols such as * or superscripts for notes or explanations
* use [note] with numbers for notes


### Basic table

#### Table 1: Mid Ulster had the highest proportion of under 25s {#table1 .tabset} 

<strong>Select the tab to view population of young or elderly people.</strong>

##### 1a Population of young people by LGD
 
```{r, echo = TRUE, warnings = FALSE, message = FALSE, out.width="100%"}
# kable code needs to be created as a string to allow correct rendering of
# source code on second tab

kable(df_mye_latest_year_youthrate,

  # left align text, right align numbers
  align = ("lrrr"),

  # full and descriptive headings
  col.names = c(
    "Local Government District (LGD)",
    "Young people in LGD",
    "Total population of LGD",
    "Young people as percentage"
  ),

  # use caption to keep text associated with table
  caption = paste0(
    "Population of young people by Local Government District in NI, ",
    latest_year
  ),

  # add thousand separator
  format.args = list(big.mark = ","),
  label = NA
) |>
  kable_styling()
```
Commentary describing the key messages you want readers to take from the table
could go here.

##### 1b Population of elderly people by LGD

```{r table_1, echo = TRUE, warnings = FALSE, message = FALSE, out.width="100%"}
# kable code needs to be created as a string to allow correct rendering of
# source code on fourth tab

kable(df_mye_latest_year_elderlyrate,

  # left align text, right align numbers
  align = ("lrrr"),

  # full and descriptive headings
  col.names = c(
    "Local Government District (LGD)",
    "Elderly people in LGD",
    "Total population of LGD",
    "Elderly people as percentage"
  ),

  # use caption to keep text associated with table
  caption = paste0(
    "Population of elderly people by Local Government District in NI, ",
    latest_year
  ),

  # add thousand separator
  format.args = list(big.mark = ","),
  label = NA
) |>
  kable_styling()
```

The key messages you want the readers to take from the table could go here.

### Pagination

#### Table 2: Northern Ireland population is growing year on year (h4)

```{r table_2, echo=TRUE, warning=FALSE, message=FALSE, out.width="100%"}
datatable(df_mye_year_gender_t,
  # DT creates row numbers by default. Suppress to ensure all cols have a header
  rownames = FALSE,
  # Rename column headers to make them descriptive
  colnames = c(
    "Mid Year Ending",
    "Females",
    "Males",
    "NI Population Total"
  ),
  caption = paste0(
    "Number of males and females making up the total NI population from ",
    earliest_year,
    "-",
    latest_year
  ),
  # left align first column and add thousand separator to numbers
  options = list(
    columnDefs = list(
      list(
        className = "dt-left",
        targets = 0
      )
    )
  )
) |>
  formatRound(
    c(
      "Females",
      "Males",
      "All persons"
    ),
    digits = 0,
    interval = 3,
    mark = ","
  )
```

This is descriptive commentary to highlight the key points in the table above.
There are more females than males year on year. Population growth is driven by
both sexes.

### Cross-tabulation
#### Table 3: Under 25s were the largest age group in Northern Ireland 

```{r table_3, echo = TRUE, warnings = FALSE, message = FALSE, out.width="100%"}
cross_tab_data <- df_mye_latest_year_agegrp_gend |>
  pivot_wider(
    names_from = gender,
    values_from = pop_total
  ) |>
  mutate(
    persons = Males + Females
  )

cross_tab_data["Female_Male_ratio_100"] <-
  round_half_up(
    cross_tab_data$Females / cross_tab_data$Males * 100, 0
  )

# Adding thousand separator
cross_tab_data$Females <- prettyNum(
  cross_tab_data$Females,
  big.mark = ","
)
cross_tab_data$Males <- prettyNum(
  cross_tab_data$Males,
  big.mark = ","
)
cross_tab_data$persons <- prettyNum(
  cross_tab_data$persons,
  big.mark = ","
)

kable(
  cross_tab_data,
  # left row labels, right align numbers
  align = "lrrrr",
  # make column headings clear and descriptive
  col.names = c(
    "Age group",
    "Females",
    "Males",
    "Total persons",
    "Number of females to every 100 males"
  ),
  # use caption for descriptive title
  caption = paste0(
    "Northern Ireland population by broad age band and sex ",
    latest_year
  ),
) |>
  kable_styling()
```

If we group the population into broad age bands we can compare the proportion of
males to females. Males outnumber females in the under 25 age group only. As age
increases, the ratio of females to males increases.

### Table with colour scales
#### Table 4: Belfast saw the greatest increase in population from `r latest_year - 1` to `r latest_year`

```{r table_4, echo = TRUE, warnings = FALSE, message = FALSE, out.width="100%"}
# function to add thousand separator
mycomma <- function(digits = 0) {
  require(formattable)
  formatter(
    "span",
    x ~ comma(
      x,
      digits = digits
    )
  )
}

# up and down arrows added to column four
# colour tiles added to column five. Colour selected is a lighter shade of
# nisra_blue (which failed accessibility check). Seems that colour specified in
# code is the mid point of the tiles generated so needs to be pale to ensure
# contrast for higher levels is sufficient.
t4 <- formattable(
  df_t4_html,
  align = c("lrrrr"),
  list(
    area(col = 2:3) ~ mycomma(digits = 0), # apply mycomma fn
    area(col = 5) ~ color_tile("transparent", "#7ca6da"),
    `Change` = formatter(
      "span",
      style = ~ style(
        color = ifelse(
          `Change` > 0,
          "black",
          "#e60000"
        )
      ),
      ~ icontext(
        ifelse(
          `Change` > 0,
          "arrow-up",
          "arrow-down"
        ),
        format(
          `Change`,
          big.mark = ","
        )
      )
    )
  )
)

t4
```

Ensure sufficient contrast between background colours and the text when using
colour scales.
