---
title: "05 Charts Sub Report"
output: html_document
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_dir = "../demo/demo_outputs/",
    output_file = paste0(
      xfun::sans_ext(input),
      "_",
      format(Sys.time(), "%d-%m-%Y_%H%M"),
      ".html"
    ),
    envir = globalenv()
  )
})
---

```{r charts_settings, include=FALSE}
# Set chunk options for all chunks in this file.

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```


```{r setup_charts}
library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

This section includes different types of charts. These include line charts,
column charts, pie charts, tree maps and a donut chart.

### Good practice for charts

* Give charts a primary title describing the main trend or message (see examples
below).
* Add a statistical subtitle that explains the measure, location and time
period. 
* Clearly label data and categories.
* Show all chart text horizontally, including axis labels.
* Write labels in full instead of using abbreviations.
* If you need to use abbreviations, explain them the first time you use them.
* Avoid using a legend where possible.
* If you need a legend, place it above the chart.
* Avoid using charts with two axes.

### Accessible charts
#### Colours

* Do not use colour on its own to show information (e.g. to read a legend).
* Make sure chart colours are easy to tell apart and easy to see against the
background.
* There is an accessible colour palette on the [Dissemination channel on NISRA Teams](https://teams.microsoft.com/l/entity/0ae35b36-0fd7-422e-805b-d53af1579093/_djb2_msteams_prefix_3480381164?context=%7B%22channelId%22%3A%2219%3A81967af3c68f424ab44a24e05c00fc6d%40thread.tacv2%22%7D&tenantId=e7a13aea-9437-4db7-a22b-cfaa4ce33b6e)
* Check contrast at least 3:1 for chart elements.
* Check contrast is at least 4.5:1 for text.

#### Help users access chart data and key messages

* Explain the main message(s) of the chart in text above or below it (i.e.
provide a clear text alternative).
* Provide the chart data in CSV or Excel format, or give users a way to request
it.
* Add alt text to static chart images describing the chart type and content.
* Think about how a screen reader will read labels and punctuation if your
charts are interactive.
* Make sure charts work well on different screen sizes, such as mobile phones,
laptops and desktop computers.
* Be aware of how interactive (e.g. Plotly) charts resize:
  * annotations easily get crowded and overlap.
  * plotly resizes the x axis which alters how the trend is perceived.
* Consider adding a message for mobile users to turn their phone sideways to
view the chart.
* If you create charts with the Plotly package, turn off the Plotly logo and 
icons. This helps reduce auditory clutter and avoids unnecessary repeated links.

For more information, see the [UK Government Analysis Function (GAF) guidance on designing charts](https://analysisfunction.civilservice.gov.uk/policy-store/data-visualisation-charts/).

### Single line chart

This section shows a static line chart and an interactive line chart. 

#### Chart features
* Text (including labels) and lines all have good colour contrasts against
white background.
* Data labels are at the end of the line to prevent overlap when the chart
changes size on different devices.

Add clear descriptive text before or after the chart. This helps all users
understand the main message of the chart.

#### Non-zero axes
To add a broken axis symbol to a plotly chart, use the "fn_break_axis" function.
Set the linecolor to match the colour used for the "y-axis" in the layout
settings. This function is a NISRA created function that can be found in the
functions folder. It can only be used with plotly charts.

To add a broken axis symbol to a 'ggplot' chart, add it as an annotation. The
'ggplot' package does not provide a built-in broken feature.

The code below shows how to add a broken axis symbol to both 'plotly' and
'ggplot' charts.


#### Figure 1: Mid year population estimates continue to increase 
##### Northern Ireland mid year population estimate (total) between `r earliest_year` and `r latest_year`

##### <span class="sr-only">Figure 1 tabs</span> {.tabset}

###### Figure 1 plotly

```{r figure_1_plotly, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 1 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%", out.height="100%"}
year1_lab <- list(
  x = ~ mid_year_ending[1] + 0.5,
  y = ~ ni_pop_total[1], # sets position of label
  xanchor = "right",
  yanchor = "middle", # alignment of label
  text = ~ paste0(
    earliest_year,
    ": \n",
    round_half_up(
      ni_pop_total[1] / 1e6,
      2
    ),
    "m"
  ),
  # punctuated and formatted label text
  showarrow = FALSE, # turn off otherwise get line between label and chart
  font = list(
    size = 14,
    color = nisra_blue
  )
)

year2_lab <- list(
  x = ~ tail(
    mid_year_ending,
    n = 1
  ), # position at last value of x variable
  y = ~ tail(
    ni_pop_total,
    n = 1
  ), # position at last value of y variable
  xanchor = "left",
  yanchor = "middle", # alignment
  # punctuated and formatted label text
  text = ~ paste0(
    latest_year,
    ": \n",
    round_half_up(
      tail(
        ni_pop_total, 1
      ) / 1e6,
      2
    ),
    "m"
  ),
  showarrow = FALSE,
  font = list(
    size = 14,
    color = nisra_blue
  )
)
# create annotations
# annotation on vertical line
note_1 <- list(
  x = 0.41,
  y = 1,
  text = "2009",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "left",
  align = "left",
  font = list(
    size = 14,
    color = nisra_navy,
    family = "Arial black"
  )
)
# use annotation rather than y axis title variable for horizontal text
yaxistitle <- list(
  x = 0.02,
  y = 1.03,
  text = "Population (Millions)",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "middle",
  xshift = 0,
  yshift = 0,
  font = list(
    size = 14,
    color = "black"
  )
)
# Create the caption - Note the non-zero axis
note_2 <- list(
  x = 0.75,
  y = -0.115,
  text = "Note the non-zero axis",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "left",
  align = "left",
  font = list(
    size = 14,
    color = nisra_navy
  )
)
# aesthetic definitions for vertical line used in fig1
vline <- function(x = 0, color = nisra_navy) {
  list(
    type = "line",
    y0 = 0,
    y1 = 1,
    yref = "paper",
    x0 = x,
    x1 = x,
    line = list(
      color = color,
      dash = "dot"
    )
  )
}

# create line chart (type = scatter and mode = line)
fig1 <- plot_ly(
  df_annual_pop_tot,
  x = ~mid_year_ending,
  y = ~ni_pop_total,
  name = "Northern Ireland population",
  line = list(
    color = nisra_blue
  ),
  type = "scatter",
  mode = "line",
  hoverinfo = "text",
  text = ~ paste0(
    mid_year_ending,
    ": ",
    format(
      round_half_up(
        ni_pop_total / 1e6,
        2
      ),
      nsmall = 2
    ),
    "m"
  )
) |>
  layout(
    showlegend = FALSE,
    font = list(
      family = "Arial",
      size = 14
    ),
    xaxis = list(
      title = "Year",
      showline = TRUE,
      linecolor = "#000000",
      showgrid = FALSE,
      showticklabels = TRUE,
      range = c(
        1999,
        2024
      ),
      ticks = "outside",
      tickwidth = 2
    ),
    yaxis = list(
      title = "",
      showline = TRUE,
      linecolor = "#000000",
      showgrid = FALSE
    ),
    annotations = list(
      year1_lab,
      year2_lab,
      note_1,
      yaxistitle,
      note_2
    )
  ) |>
  # To add the broken axis symbol pipe the function "fn_break_axis" with
  # "linecolor" matching the colour given above. This function is a NISRA
  # created function that can be found in the functions folder.
  fn_break_axis(
    linecolor = "#000000",
    ref_line = vline(2009)
  ) |>
  f_responsive_annotations()
# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig1 <- fig1 |>
  config(
    displayModeBar = FALSE
  )

fig1 <- htmlwidgets::onRender(
  fig1,
  "
  function(el, x) {
    window.fig1_chart = el;

    setTimeout(function() {
      Plotly.Plots.resize(el);
    }, 200);
  }
  "
)

fig1
```

```{r figure1_dl_buttons_plotly}
# call make tables function to create chart download buttons
f_make_tables(
  data = df_fig1_xls,
  title = fig1_title,
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig1_chart"
)
```

###### Figure 1 plotly zero axis

```{r figure_1_plotly_zero_axis, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 1 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%", out.height="100%"}
year1_lab <- list(
  x = ~ mid_year_ending[1],
  y = ~ ni_pop_total[1], # sets position of label
  xanchor = "right",
  yanchor = "middle", # alignment of label
  text = ~ paste0(
    earliest_year,
    ": \n",
    round_half_up(
      ni_pop_total[1] / 1e6,
      2
    ),
    "m"
  ),
  # punctuated and formatted label text
  showarrow = FALSE, # turn off otherwise get line between label and chart
  font = list(
    size = 14,
    color = nisra_blue
  )
)

year2_lab <- list(
  x = ~ tail(
    mid_year_ending,
    n = 1
  ), # position at last value of x variable
  y = ~ tail(
    ni_pop_total,
    n = 1
  ), # position at last value of y variable
  xanchor = "left",
  yanchor = "middle", # alignment
  # punctuated and formatted label text
  text = ~ paste0(
    latest_year,
    ": \n",
    round_half_up(
      tail(ni_pop_total, 1) / 1e6,
      2
    ),
    "m"
  ),
  showarrow = FALSE,
  font = list(
    size = 14,
    color = nisra_blue
  )
)

# create annotations
# annotation on vertical line
note_1 <- list(
  x = 0.41,
  y = 1, # position of annotation
  text = "2009",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "left",
  align = "left",
  font = list(
    size = 14,
    color = nisra_navy,
    family = "Arial black"
  )
)

# use annotation rather than y axis title variable for horizontal text
yaxistitle <- list(
  x = 0.02,
  y = 1.06,
  text = "Population (Millions)",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "centre",
  xshift = 0,
  yshift = 0,
  font = list(
    size = 14,
    color = "black"
  )
)

# aesthetic definitions for vertical line used in fig1
vline <- function(x = 0, color = nisra_navy) {
  list(
    type = "line",
    y0 = 0,
    y1 = 1,
    yref = "paper",
    x0 = x,
    x1 = x,
    line = list(
      color = color,
      dash = "dot"
    )
  )
}

# create line chart (type = scatter and mode = line)
fig1 <- plot_ly(
  df_annual_pop_tot,
  x = ~mid_year_ending,
  y = ~ni_pop_total,
  name = "Northern Ireland population",
  line = list(
    color = nisra_blue
  ),
  type = "scatter",
  mode = "line",
  hoverinfo = "text",
  # formatting hover text to millions(m) to 2dp
  text = ~ paste0(
    mid_year_ending,
    ": ",
    format(
      round_half_up(
        ni_pop_total / 1e6,
        2
      ),
      nsmall = 2
    ),
    "m"
  )
) |>
  layout(
    showlegend = FALSE,
    font = list(
      family = "Arial",
      size = 14
    ),
    xaxis = list(
      title = "Year",
      showline = TRUE,
      linecolor = "#000000",
      showgrid = FALSE,
      showticklabels = TRUE,
      range = c(
        1999,
        2024
      ),
      ticks = "outside",
      tickwidth = 2
    ),
    yaxis = list(
      title = "",
      showline = TRUE,
      range = c(
        0,
        2100000
      ),
      linecolor = "#000000",
      showgrid = FALSE
    ),
    annotations = list(
      year1_lab,
      year2_lab,
      note_1,
      yaxistitle
    )
  ) |>
  layout(
    shapes = list(
      vline(2009)
    )
  ) |>
  f_responsive_annotations()

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig1 <- fig1 |>
  config(
    displayModeBar = FALSE
  )

fig1 <- htmlwidgets::onRender(
  fig1,
  "
  function(el, x) {
    window.fig1_zero_chart = el;
  }
  "
)

fig1
```

```{r figure1_dl_buttons_plotly_zero_axis}
# call make tables function to create chart download buttons
f_make_tables(
  data = df_fig1_xls,
  title = fig1_title,
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig1_zero_chart"
)
```

###### Figure 1 static 

```{r figure_1_static, echo = TRUE, out.width="100%", fig.alt="line chart showing growth in population from 2001 to 2019. More detail on trends is included in text immediately above or below the chart" }
fig1_static <- ggline(
  df_annual_pop_tot,
  x = "mid_year_ending",
  y = "ni_pop_total",
  plot_type = "l",
  color = nisra_blue,
  linetype = "solid",
  size = 1,
  caption = "Note the non-zero axis",
  xlim = c(
    2000,
    2026
  ),
  xlab = "Mid year ending",
  ylab = "Population"
) +
  font(
    "ylab",
    angle = 0,
    vjust = 1.02
  ) +
  scale_y_continuous(
    limits = c(
      1650000,
      2000000
    ),
    labels = scales::label_number(
      suffix = "m",
      scale = 1e-6,
      big.mark = ","
    )
  ) +
  theme(
    text = element_text(
      family = "Arial"
    )
  ) +
  theme(
    axis.title = element_text(
      size = 11
    ),
    axis.text = element_text(
      size = 10
    ),
    # give a bit of room at the left for the slashes if needed
    plot.margin = margin(
      5.5,
      20,
      5.5,
      20
    )
  ) +
  geom_text(
    data = df_annual_pop_tot |>
      filter(
        mid_year_ending == latest_year
      ),
    aes(
      label = paste0(
        latest_year,
        ": \n",
        format(
          ni_pop_total,
          big.mark = ","
        )
      ),
      x = mid_year_ending + 1.5,
      y = ni_pop_total
    )
  ) +
  geom_text(
    data = df_annual_pop_tot |>
      filter(
        mid_year_ending == earliest_year
      ),
    aes(
      label = paste0(
        earliest_year,
        ": ",
        format(
          ni_pop_total,
          big.mark = ","
        )
      ),
      x = mid_year_ending + 2.5,
      y = ni_pop_total - 7000
    )
  ) +
  geom_vline(
    xintercept = 2009,
    linetype = "dashed",
    color = nisra_blue
  ) +
  annotate(
    "text",
    x = 2009.5,
    y = 1880000,
    hjust = 0,
    label = paste(
      strwrap(
        "2009: Growth rate starts to slow",
        width = 18
      ),
      collapse = "\n"
    )
  ) +
  # --- Add a “fake” axis-break mark (two slashes) near the left axis ---
  coord_cartesian(
    clip = "off"
  ) +
  annotate(
    "text",
    x = min(
      df_annual_pop_tot$mid_year_ending
    ),
    y = 1.67e6,
    label = "//",
    hjust = 2.55, # nudges further left
    vjust = 0.5,
    size = 5
  )

fig1_static
```

```{r figure1_dl_buttons_static}
# call make tables function to create chart download buttons
f_make_tables(
  data = df_fig1_xls,
  title = fig1_title,
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  ggplot = TRUE
)
```

##### <span class="sr-only">End tabs</span>

Further information on the chart download buttons can be found in
section [7. Download data](#download-data) of the demo report.

### Multiple lines chart
This section shows a line chart with more than one line. The lines are easy to
see because they do not overlap.

#### Features
* The chart uses good contrast (different shades of the same colour) to
distinguish between males and females.
* The legend is placed above the chart.
* A static chart is used here so it views well on mobile as well as desktop.

Include descriptive text before or after the chart.


#### Figure 2: Females outnumber males in total population

##### Northern Ireland mid year population estimate (sex) `r earliest_year` to `r latest_year` 

```{r line_chart_multiple, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 2 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
# Create separate dataframe for selected datapoints from main dataframe
df_mye_selected_years_gender <- df_mye_year_gender |>
  filter(
    mid_year_ending %in% c(
      2001,
      2012,
      2022
    )
  ) |>
  # format numbers
  mutate(
    label = paste0(
      mid_year_ending,
      ": ",
      round_half_up(
        mye_pop / 1e6,
        2
      ),
      "m"
    )
  )


# code to create interpolated points with blank labels along the line of the
# line plot to cause repulsion of labels from line and avoid overlap - the
# final df, df_fig2_labs_inter is used in the geom_text_repel statement

df_mye_interp_male <- filter(
  df_mye_year_gender,
  gender == "Males"
)

df_mye_interp_male <- as.data.frame(
  do.call(
    cbind,
    approx(
      df_mye_interp_male$mid_year_ending,
      df_mye_interp_male$mye_pop,
      n = 30
    )
  )
)

df_mye_interp_female <- filter(
  df_mye_year_gender,
  gender == "Females"
)

df_mye_interp_female <- as.data.frame(
  do.call(
    cbind,
    approx(
      df_mye_interp_female$mid_year_ending,
      df_mye_interp_female$mye_pop,
      n = 30
    )
  )
)

df_mye_interp_male$label <- ""
df_mye_interp_male$Gender <- "Males"

df_mye_interp_male <- df_mye_interp_male |>
  select(
    x,
    Gender,
    y,
    label
  )

colnames(df_mye_interp_male) <- colnames(
  df_mye_selected_years_gender
)

df_mye_lab_select_male <- filter(
  df_mye_selected_years_gender,
  gender == "Males"
)

df_mye_interp_male <- rbind(
  df_mye_lab_select_male,
  df_mye_interp_male
)

df_mye_interp_female$label <- ""
df_mye_interp_female$Gender <- "Females"

df_mye_interp_female <- df_mye_interp_female |>
  select(
    x,
    Gender,
    y,
    label
  )

colnames(df_mye_interp_female) <- colnames(
  df_mye_selected_years_gender
)

df_mye_lab_select_female <- filter(
  df_mye_selected_years_gender,
  gender == "Females"
)

df_mye_interp_female <- rbind(
  df_mye_lab_select_female,
  df_mye_interp_female
)

df_fig2_labs_inter <- rbind(
  df_mye_interp_male,
  df_mye_interp_female
)

df_fig2_labs_inter <- arrange(
  df_fig2_labs_inter,
  mid_year_ending,
  gender
)


fig2 <- ggplot(
  df_mye_year_gender,
  aes(
    x = mid_year_ending,
    y = mye_pop,
    color = gender
  )
) +
  geom_line(
    linewidth = 1
  ) +
  # ACC: Change to line colours with sufficient contrast
  scale_color_manual(
    values = c(
      nisra_blue,
      nisra_navy
    )
  ) +

  # Using NITotals_lab_select_multiple from above to pick the years that will
  # have "points", specifying the size of the dots
  geom_point(
    data = df_mye_selected_years_gender,
    aes(
      x = mid_year_ending,
      y = mye_pop
    ),
    size = 3
  ) +
  labs(
    x = "Year",
    y = "Population",
    caption = "Note the non-zero axis"
  ) +
  theme_bw() +
  theme(
    # removes grid lines from chart
    panel.grid = element_blank(),
    # removes chart border top and right
    panel.border = element_blank(),
    axis.line.x = element_line(
      colour = "#000000"
    ),
    axis.line.y = element_line(
      colour = "#000000"
    ),
    # ACC: make y axis title horizontal and position it further from y axis line
    axis.title.y = element_text(
      angle = 0,
      vjust = 1,
      hjust = 0,
      margin = margin(
        t = 0,
        r = 5,
        b = 0,
        l = 0
      )
    ),
    # align title and subtitle to left
    plot.title.position = "plot",
    # align caption to left
    plot.caption = element_text(
      hjust = 0
    ),
    # ACC: change all fonts to 12
    text = element_text(
      size = 12,
      family = "Arial"
    ),
    # bolds title and subtitle + adds margin below subtitle and plot
    plot.title = element_text(
      face = "bold"
    ),
    plot.subtitle = element_text(
      face = "bold",
      margin = margin(
        0,
        0,
        15,
        0
      )
    ),
    # ACC: Legend to top and tidy up
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 12
    )
  ) +
  geom_text_repel(
    # ACC: show value label for specified years
    data = df_fig2_labs_inter,
    aes(
      x = mid_year_ending,
      y = mye_pop,
      label = label
    ),
    force = 10
  ) +
  # ACC: make x axis labels easier to read by removing every other year
  scale_x_continuous(
    breaks = seq(
      earliest_year,
      latest_year,
      by = 2
    )
  ) +
  # add comma to y labels
  scale_y_continuous(
    labels = label_number(
      scale = 1e0,
      big.mark = ","
    )
  ) +
  # indicate non zero axis
  labs(
    tag = "//"
  ) +
  theme(
    plot.tag.position = c(
      0.203,
      0.2
    )
  )

fig2
```

```{r figure_2_dl_buttons}
f_make_tables(
  data = df_fig2_xls,
  title = paste0(
    "Figure 2: Northern Ireland mid year population ",
    "estimate (gender) ",
    earliest_year,
    " to ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  ggplot = TRUE
)
```

### Multiple single line charts

This section shows a different way to display data that would otherwise have too
many lines. The Government Analysis Function recommends aiming for no more than
4 lines.

#### Features
* These 'small multiples' (sometimes called 'panel charts') are an option for
more than 4 lines. 
* The y axis is identical on all charts.
* This chart is static because it works well on mobiles and when magnified.

You could also highlight one line in colour and make all the other lines grey if
that shows your message more clearly.

Add clear text before or after the chart to explain the main message.

#### Figure 3: In `r latest_year`, Belfast Local Government District (LGD) has more younger people than other areas

##### Northern Ireland mid year population estimate by LGD in `r latest_year`

```{r line_chart_facet, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 3 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
fig3 <- ggplot(
  df_mye_latest_5yr_age_lgd,
  aes(
    x = age_group_5,
    y = population_latest
  )
) +
  geom_line(
    linewidth = 1,
    group = 1,
    color = nisra_blue
  ) +
  # ACC: Specify line colour with sufficient contrast
  facet_wrap(
    vars(
      str_wrap(
        lgd2014name,
        30
      )
    ),
    ncol = 3
  ) +
  # ACC: Use clear and descriptive labels
  labs(
    x = "Age",
    y = "Population"
  ) +
  theme_bw() +
  theme(
    # removes grid lines from chart
    panel.grid = element_blank(),
    # panel margin
    panel.spacing.x = unit(
      0.7,
      "lines"
    ),
    axis.line.x = element_line(
      colour = "#000000"
    ),
    axis.line.y = element_line(
      colour = "#000000"
    ),
    # ACC: make y axis title horizontal and position it further from y axis line
    axis.title.y = element_text(
      angle = 1,
      vjust = 1,
      hjust = 0,
      margin = margin(
        t = 0,
        r = 5,
        b = 0,
        l = 0
      )
    ),
    # ACC: change all fonts to 12
    text = element_text(
      size = 12,
      family = "Arial"
    ),
    # facet only legend hide bg
    legend.title = element_blank(),
    legend.position = "none",
    strip.background = element_blank()
  ) +
  scale_x_discrete(
    breaks = c(
      "0-4",
      "45-49",
      "90-94"
    ),
    labels = c(
      "0-4",
      "45-49",
      "90+"
    )
  ) +
  # ACC: add comma to y labels
  scale_y_continuous(
    labels = label_number(
      scale = 1e0,
      big.mark = ","
    )
  )

fig3
```

```{r figure_3_dl_buttons}
# download buttons
# download buttons
f_make_tables(
  data = df_fig3_xls,
  title = paste0(
    "Figure 3: Northern Ireland mid year population estimate by 5 ",
    "year age groups and LGD in ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  ggplot = TRUE
)
```

### Bar chart (a-z)

This section presents two versions of a bar chart: a static chart produced using
ggplot and an interactive chart produced using ggplot and ggplotly. The chart is
ordered alphabetically by Local Government District (LGD). A bar chart was
selected to allow the full LGD names to be displayed clearly along the
horizontal axis. Please note that interactive charts may not show everything
when zoomed in.

#### Features

* Long council names are displayed in full without wrapping or having text at an
angle with this orientation of bars.

Include descriptive text before or after the chart.

#### Figure 4: Young people by LGD

##### Data from `r latest_year` mid-year estimates (MYEs) showing the percentage of each LGD population that is under 25

##### <span class="sr-only">Figure 4 tabs</span> {.tabset}

###### Figure 4 static (ggplot)

```{r bar_chart_az, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 4 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
# create amended df for chart. Reorder factors for plotting (need lgd names in
# reverse alphabetical order as coord flip reverses order).
# Remove unnecessary columns
fig4_data <- df_mye_latest_year_youthrate |>
  mutate(
    lgd2014name = fct_reorder(
      lgd2014name,
      desc(lgd2014name)
    )
  ) |>
  select(
    -under25_pop,
    -lgd_pop_total
  )

ggplot(
  fig4_data,
  aes(
    x = lgd2014name,
    y = youthrate
  )
) +
  geom_bar(
    fill = nisra_blue,
    stat = "identity"
  ) +
  geom_text(
    aes(
      label = youthrate
    ),
    hjust = 1.5,
    colour = "white",
    fontface = "bold"
  ) +
  # adds value labels to the bars, sets text to colour and weight that meet
  # contrast standards.
  theme_bw() +
  # axis labels
  labs(
    x = paste(
      strwrap(
        "LGD",
        width = 0.2 * getOption("width")
      ),
      collapse = "\n"
    ),
    y = "Percentage of population under 25"
  ) +
  # makes y axis title horizontal
  theme(
    axis.title.y = element_text(
      angle = 0,
      vjust = 1,
      hjust = 1
    ),
    # removes grids from chart
    panel.grid = element_blank(),
    # added vertical grid at major ticks
    panel.grid.major.x = element_line(
      color = "#BEBEBE",
      linewidth = 0.1
    ),
    # removes chart border top and right
    panel.border = element_blank(),
    # xaxis line black
    axis.line.x = element_line(
      colour = "#000000"
    ),
    # yaxis line black
    axis.line.y = element_line(
      colour = "#000000"
    )
  ) +
  coord_flip()
```

###### Figure 4 interactive (ggplotly)

```{r bar_chart_az_ggploty, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 4 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
# Create amended data frame for chart.
# Reorder factors for plotting, as coord_flip() reverses the order.
# Remove unnecessary columns.
fig4_data <- df_mye_latest_year_youthrate |>
  mutate(
    lgd2014name = fct_reorder(
      lgd2014name,
      desc(lgd2014name)
    )
  ) |>
  select(
    -under25_pop,
    -lgd_pop_total
  )

# Create static ggplot
fig4 <- ggplot(
  fig4_data,
  aes(
    x = lgd2014name,
    y = youthrate,
    text = paste0(
      "<b>",
      lgd2014name,
      "</b>",
      "<br>Percentage of population under 25: ",
      sprintf(
        "%.1f",
        youthrate
      )
    )
  )
) +
  geom_bar(
    fill = nisra_blue,
    stat = "identity"
  ) +
  # Set labels for bars
  geom_text(
    aes(
      label = sprintf(
        "%.1f",
        youthrate
      )
    ),
    hjust = 0.9,
    colour = "white",
    fontface = "bold"
  ) +
  theme_bw() +
  labs(
    x = paste(
      strwrap(
        "LGD",
        width = 0.2 * getOption("width")
      ),
      collapse = "\n"
    ),
    y = "Percentage of population under 25"
  ) +
  theme(
    axis.title.y = element_text(
      angle = 0,
      vjust = 1,
      hjust = 0
    ),
    panel.grid = element_blank(),
    panel.grid.major.x = element_line(
      colour = "#BEBEBE",
      linewidth = 0.1
    ),
    panel.border = element_blank(),
    axis.line.x = element_line(
      colour = "#000000"
    ),
    axis.line.y = element_line(
      colour = "#000000"
    )
  ) +
  coord_flip()

# Convert to interactive plotly chart
fig4_interactive <- ggplotly(
  fig4,
  tooltip = "text"
) |>
  layout(
    yaxis = list(
      title = ""
    ),
    # Creates an annotation for the 'LGD' y axis label
    annotations = list(
      list(
        x = -0.1,
        y = 1.01,
        xref = "paper",
        yref = "paper",
        text = "<b>LGD</b>",
        showarrow = FALSE,
        xanchor = "right",
        yanchor = "bottom",
        font = list(
          size = 12
        )
      )
    ),
    margin = list(
      l = 180,
      r = 30,
      t = 40,
      b = 40
    ),
    hoverlabel = list(
      align = "left"
    )
  ) |>
  config(
    displayModeBar = FALSE
  )

# Move the value labels inside the bars
fig4_interactive$x$data[[2]]$x <-
  fig4_interactive$x$data[[2]]$x - 2

fig4_interactive <- htmlwidgets::onRender(
  fig4_interactive,
  "
  function(el, x) {
    window.fig4_chart = el;
  }
  "
)

fig4_interactive
```

##### <span class="sr-only">End tabs</span>

```{r figure_4_dl_buttons}
# download buttons
f_make_tables(
  data = df_fig4_5_xls,
  title = "Figure 4: Percentage of young people by LGD",
  data_style = ns_percentage,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig4_chart"
)
```

The chart shows that the LGD with the greatest percentage of population less
than 25 is Mid Ulster. Ards and North Down is the LGD with the smallest
percentage.

### Interactive bar chart using plotly
This bar chart is created using the same data but using plotly to create an
interactive html chart. When resized for mobile the bars shorten but the labels
remain a more readable size. Keep the x axis title short so it won't get cut off
on a smaller display.

Include descriptive text before or after the chart highlighting the key message.

#### Figure 5: Proportion of under 25s was higher in `r earliest_year`.

##### Percentage of population aged under 25 by LGD, `r earliest_year` MYEs

Dashed line represents Northern Ireland mean.

```{r bar_chart_interactive, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 5 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
# create annotation for y axis title so it can be horizontal.
yaxistitle <- list(
  x = -0.1,
  y = 1.01, # position of y axis title
  text = "LGD", # title
  showarrow = FALSE,
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "middle", # alignment
  xshift = 0,
  yshift = 0,
  font = list(
    family = "Arial Black",
    size = 14,
    color = "black"
  )
)

fig5 <- plot_ly(
  df_mye_earliest_youthrate,
  x = ~earliest_youthrate,
  y = ~lgd2014name,
  type = "bar",
  orientation = "h", # horizontal bars. Default is vertical
  hoverinfo = "text",
  # adds value labels to the bars and sets hover text
  text = ~earliest_youthrate,
  textposition = "auto",
  # sets colour of bars to navy
  marker = list(
    color = nisra_navy
  )
) |>
  layout(
    # add annotation
    font = list(
      family = "Arial",
      size = 14
    ),
    annotations = list(
      yaxistitle
    ),
    shapes = list(
      vline(
        ~ mean(earliest_youthrate),
        color = nisra_blue
      )
    ),
    xaxis = list(
      title = "Under 25s (%)"
    ),
    yaxis = list(
      title = "",
      ticks = "outside",
      tickcolor = "#ffffff",
      ticklen = 10,
      categoryorder = "category descending"
    )
  ) |>
  f_responsive_annotations()

# displays lgds in alphabetic order
# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.

fig5 <- fig5 |>
  config(
    displayModeBar = FALSE,
    displaylogo = FALSE
  )

fig5 <- htmlwidgets::onRender(
  fig5,
  "
  function(el, x) {
    window.fig5_chart = el;
  }
  "
)

fig5
```

```{r figure_5_dl_buttons}
# download buttons
f_make_tables(
  df_fig4_5_xls,
  title = paste0(
    "Figure 5: Percentage of under 25s by LGD (",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig5_chart"
)
```

The chart shows that LGD with the greatest percentage of population under 25 in
`r earliest_year` was Derry City and Strabane. Ards and North Down was the LGD
with the smallest percentage.

### Column chart 

This section includes two versions of a column chart (one static, one
interactive) and two stacked bar charts.

#### Features

* The contrasting shades of blue make it easy to differentiate between male and
female.
* Data labels have been added.
* The order of the legend is the same as the column order.
* The plotly x axis labels are short and don't overlap when resized.
* The plotly data labels will change to vertical on mobile which is not ideal
but the hover labels will be horizontal and legible.
* The plotly stacked bar age group labels are short and resize well on mobile.
* The horizontal stacked bar resizes well because there are only 2 groups. If
there were many more a static chart would likely be more appropriate.

Include descriptive text before or after the chart.

#### Figure 6: Females outnumber males in all age groups except the under 25s

##### Northern Ireland population in `r latest_year` by age group

##### <span class="sr-only">Figure 6 tabs</span> {.tabset}

###### ggplot static 

```{r column_chart, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 6 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below.", out.width="100%"}
fig6_static <- ggplot(
  df_mye_latest_year_agegrp_gend,
  aes(
    x = age_group,
    y = pop_total,
    fill = gender
  )
) +
  geom_col(
    position = position_dodge()
  ) +
  # colour bars that meet contrast for adjacent colours
  scale_fill_manual(
    values = c(
      nisra_navy,
      nisra_blue
    )
  ) +
  # add value labels, and add thousand separator
  geom_text(
    aes(
      label = prettyNum(
        pop_total,
        big.mark = ","
      )
    ),
    position = position_dodge(
      width = 1
    ),
    vjust = 1.1,
    colour = "white",
    size = 3.2
  ) +
  theme_bw() +
  # axis labels
  labs(
    x = "Age group",
    y = "Population",
    fill = "Gender"
  ) +
  theme(
    axis.title.y = element_text(
      # makes y axis title horizontal
      angle = 0,
      vjust = 1.05,
      hjust = 1
    ),
    # removes grids from chart
    panel.grid = element_blank(),
    # added horizontal grid at major ticks
    panel.grid.major.y = element_line(
      color = "#BEBEBE",
      linewidth = 0.1
    ),
    # removes chart border top and right
    panel.border = element_blank(),
    # xaxis line black
    axis.line.x = element_line(
      colour = "#000000"
    ),
    # yaxis line black
    axis.line.y = element_line(
      colour = "#000000"
    ),
    legend.position = "top"
  ) +
  # add thousand separator to y axis labels
  scale_y_continuous(
    labels = label_number(
      scale = 1e0,
      big.mark = ","
    )
  )

fig6_static
```

```{r fig6_static_download}
f_make_tables(
  df_fig6_xls,
  title = paste0(
    "Figure 6 Static: LGD populations by sex and age group (",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  ggplot = TRUE
)
```

###### plotly 

```{r figure_6b,  echo = TRUE, fig.alt="Figure 6b alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig6_grouped <- pivot_wider(
  df_mye_latest_year_agegrp_gend,
  names_from = gender,
  values_from = pop_total
)

fig6_grouped <- as.data.frame(
  fig6_grouped
)

fig6_grouped <- pivot_longer(
  fig6_grouped,
  cols = c(
    Females,
    Males
  ),
  names_to = "variable",
  values_to = "value"
)

nisra_blue <- "#3878c5"
nisra_navy <- "#00205b"

colormap <- setNames(
  object = c(
    "#00205b",
    "#3878c5"
  ),
  nm = unique(
    fig6_grouped$variable
  )
)

fig6_grouped <- fig6_grouped |>
  mutate(
    variable = factor(
      variable,
      levels = c(
        "Females",
        "Males"
      )
    )
  ) |>
  plotly::plot_ly(
    x = ~age_group,
    y = ~value,
    color = ~variable,
    colors = colormap,
    type = "bar",
    text = format(
      fig6_grouped$value,
      big.mark = ","
    ),
    textposition = "inside",
    textangle = "horizontal", # for accessibility
    insidetextfont = list(
      size = 12
    ), # to fit
    hovertemplate = "%{text}", # hover text = text
    textfont = list(
      color = "white"
    )
  ) |>
  layout(
    # makes it a clustered bar
    barmode = "group",
    font = list(
      family = "Arial",
      size = 14
    ),
    xaxis = list(
      title = "Age group"
    ),
    yaxis = list(
      title = "",
      tickformat = ",d",
      showgrid = FALSE
    ),
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = 1.1,
      # so order is same as legend
      traceorder = "normal"
    ),
    annotations = list(
      list(
        text = "Population",
        x = -0.03,
        xref = "paper",
        xanchor = "center",
        y = 1.02,
        yref = "paper",
        yanchor = "bottom",
        showarrow = FALSE,
        font = list(
          size = 16
        )
      )
    )
  ) |>
  # modebar is series of icons and plotly logo linked to plotly site. Turned off
  # throughout to avoid multiple links going to same destination (poor
  # accessibility) and icons can create visual clutter.
  config(
    displayModeBar = FALSE
  )

fig6_grouped <- htmlwidgets::onRender(
  fig6_grouped,
  "
  function(el, x) {
    window.fig6_grouped_chart = el;
  }
  "
)

fig6_grouped
```

```{r fig6_grouped_download}
f_make_tables(
  df_fig6_xls,
  title = paste0(
    "Figure 6 Grouped: LGD populations by sex and age group (",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig6_grouped_chart"
)
```


###### Stacked bar 

```{r figure_6c, echo = TRUE, fig.alt="Figure 6e alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig6_grouped <- pivot_wider(
  df_mye_latest_year_agegrp_gend,
  names_from = gender,
  values_from = pop_total
)

fig6_grouped <- as.data.frame(
  fig6_grouped
)

fig6_grouped <- pivot_longer(
  fig6_grouped,
  cols = c(
    Females,
    Males
  ),
  names_to = "variable",
  values_to = "value"
)

fig6_stacked <- fig6_grouped |>
  mutate(
    variable = factor(
      variable,
      levels = c(
        "Females",
        "Males"
      )
    )
  ) |>
  plotly::plot_ly(
    x = ~age_group,
    y = ~value,
    color = ~variable,
    colors = colormap,
    type = "bar",
    text = format(
      fig6_grouped$value,
      big.mark = ","
    ),
    textposition = "inside",
    textangle = "horizontal", # for accessibility
    insidetextfont = list(
      size = 12
    ), # to fit
    hovertemplate = "%{text}", # hover text = text
    textfont = list(
      color = "white"
    )
  ) |>
  layout(
    # stacked rather than clustered/grouped
    barmode = "stack",
    font = list(
      family = "Arial",
      size = 14
    ),
    xaxis = list(
      title = "Age group"
    ),
    yaxis = list(
      title = "",
      tickformat = ",d",
      showgrid = FALSE
    ),
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = 1.1,
      font = list(
        family = "Arial",
        size = 14
      ),
      traceorder = "normal"
    ),
    annotations = list(
      list(
        text = "Population",
        x = -0.03,
        xref = "paper",
        xanchor = "center",
        y = 1.02,
        yref = "paper",
        yanchor = "bottom",
        showarrow = FALSE,
        font = list(
          size = 16
        )
      )
    )
  )

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig6_stacked <- config(
  fig6_stacked,
  displayModeBar = FALSE
)

fig6_stacked <- htmlwidgets::onRender(
  fig6_stacked,
  "
  function(el, x) {
    window.fig6_stacked_chart = el;
  }
  "
)

fig6_stacked
```

```{r fig6_stacked_download}
f_make_tables(
  df_fig6_xls,
  title = paste0(
    "Figure 6 Stacked: LGD populations by sex and age group (",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig6_stacked_chart"
)
```

###### Stacked horizontal bar (%) 

```{r figure_6d, echo = TRUE, fig.alt="Figure 6d alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
# for stacked horizontal bar set orientation to "h" (horizontal)
# and barmode to stack
fig6_horizontal <- plot_ly(
  df_latest_year_agegrp_gend_pct,
  x = ~Females,
  y = ~age_group,
  marker = list(
    color = nisra_navy
  ),
  type = "bar",
  orientation = "h",
  name = "Females",
  text = ~ paste0(
    format(
      round_half_up(
        Females,
        1
      ),
      nsmall = 1
    ),
    "%"
  ),
  # use excel rounding function, display to 1dp
  insidetextanchor = "middle",
  hoverinfo = "y+text"
) |>
  add_trace(
    x = ~Males,
    y = ~age_group,
    orientation = "h",
    marker = list(
      color = nisra_blue
    ),
    name = "Males",
    text = ~ paste0(
      format(
        round_half_up(
          Males,
          1
        ),
        nsmall = 1
      ),
      "%"
    )
  ) |>
  layout(
    barmode = "stack",
    font = list(
      family = "Arial",
      size = 14
    ),
    yaxis = list(
      title = "",
      ticks = "outside",
      tickcolor = "#ffffff",
      ticklen = 5,
      # default was 65+ at top
      autorange = "reversed"
    ),
    xaxis = list(
      title = "Percentage (%)"
    ),
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.42,
      y = 1.1,
      traceorder = "normal"
    ),
    annotations = list(
      list(
        text = "Age group",
        x = 0,
        xref = "paper",
        xanchor = "center",
        y = 1.02,
        yref = "paper",
        yanchor = "bottom",
        showarrow = FALSE,
        font = list(
          size = 16
        )
      )
    )
  )

# ensure legend is same order as stacks
# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig6_horizontal <- config(
  fig6_horizontal,
  displayModeBar = FALSE
)

fig6_horizontal <- htmlwidgets::onRender(
  fig6_horizontal,
  "
  function(el, x) {
    window.fig6_horizontal_chart = el;
  }
  "
)

fig6_horizontal
```

```{r fig6_horizontal_download}
f_make_tables(
  df_fig6_xls,
  title = paste0(
    "Figure 6 Horizontal: LGD populations by sex and age group (",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig6_horizontal_chart"
)
```

##### <span class="sr-only">End tabs</span>


### Column chart 5 categories

This chart has been included to show how the 5 colour palette can be used for
categorical data. 

#### Features
* The colours are used in order to ensure sufficient contrast between adjacent
bars.
* The text labels are black or white as appropriate against the background
colour.
* These colours can be used for fewer categories as long as the order is
retained.

#### Figure 7: Over 85s remain the smallest age group

##### Northern Ireland population in 2022 in 5 age bands

```{r figure_7, echo = TRUE, fig.alt="Figure 7 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
yaxistitle <- list(
  text = "Population",
  x = 0.01,
  y = 1.05,
  xref = "paper",
  yref = "paper",
  xanchor = "right",
  yanchor = "top",
  showarrow = FALSE,
  font = list(
    size = 14
  )
)

fig7 <- plot_ly(
  df_mye_latest_5_agegroups,
  x = ~age_grp_5,
  y = ~agegroup_total,
  type = "bar",
  marker = list(
    color = c(
      nisra_blue,
      nisra_navy,
      nisra_col3_green,
      nisra_col4_purple,
      nisra_col5_lilac
    )
  ),
  # order of colours is important: need medium and
  # dark colours beside one another
  text = ~ format(
    agegroup_total,
    big.mark = ","
  ),
  textfont = list(
    color = c(
      "white",
      "white",
      "black",
      "white",
      "black"
    )
  ),
  # need to specify black or white labels. plotly default
  # is grey and fails contrast required
  textangle = "horizontal",
  hovertemplate = ~ paste0(
    age_group,
    ": ",
    format(
      agegroup_total,
      big.mark = ","
    ),
    "<extra></extra>"
  ),
  hoverlabel = list(
    font = list(
      color = c(
        "white",
        "white",
        "black",
        "white",
        "black"
      )
    )
  )
) |>
  layout(
    yaxis = list(
      title = "",
      tickformat = ",d"
    ), # adds thousand separator
    xaxis = list(
      title = "Age group"
    ),
    annotations = list(
      yaxistitle
    )
  )

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig7 <- fig7 |>
  config(
    displayModeBar = FALSE
  )

fig7 <- htmlwidgets::onRender(
  fig7,
  "
  function(el, x) {
    window.fig7_chart = el;
  }
  "
)

fig7
```

```{r figure_7_dl_buttons}
f_make_tables(
  df_fig7_xls,
  title = paste0(
    "Figure 7: Northern Ireland population by 5 age bands, ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig7_chart"
)
```

### Pie chart

Good practice is to only use pie charts with 5 categories or fewer. 

#### Features

* Category labels are added to avoid using a legend.
* Value labels have been added on segments so that users do not have to estimate
relative sizes.
* These pie charts are in two columns so that when resized for mobile they will
wrap one under the other and remain readable. 

Include descriptive text before or after the chart.

#### Figure 8: Higher percentage of females than males in the over 65 age group in NI

##### Percentage of Northern Ireland population aged 65 plus by sex, `r latest_year`

```{r pie_chart, echo = TRUE, fig.alt="Figure 8 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below"}
fig8a <- plot_ly(
  df_mye_latest_over_under_65,
  labels = ~age_65split,
  values = ~Females,
  type = "pie",
  # create better punctuated labels for screen readers,
  # default value failed accessibility audit
  text = ~ paste0(
    age_65split,
    " \n (",
    round_half_up(
      Females / sum(Females) * 100,
      1
    ),
    "%)"
  ),
  # increase label size
  insidetextfont = list(
    size = 16
  ),
  insidetextorientation = "horizontal",
  textinfo = "text",
  hoverinfo = "text",
  # hover font colour for contrast
  hoverlabel = list(
    font = list(
      color = c(
        "black",
        "white"
      )
    )
  ),
  marker = list(
    colors = c(
      nisra_col3_green,
      nisra_navy
    )
  ),
  textfont = list(
    color = c(
      "black",
      "white"
    )
  ),
  showlegend = FALSE
) |>
  # modebar is series of icons and plotly logo linked to plotly site. Turned off
  # throughout to avoid multiple links going to same destination (poor
  # accessibility) and icons can create visual clutter.
  config(
    displayModeBar = FALSE
  ) |>
  layout(
    annotations = list(
      x = 0.5,
      y = 1.08,
      xanchor = "center",
      text = "<b> Females </b>",
      showarrow = FALSE,
      font = list(
        family = "Arial",
        size = 16
      )
    ),
    font = list(
      family = "Arial",
      size = 14
    )
  )

fig8a <- htmlwidgets::onRender(
  fig8a,
  "
  function(el, x) {
    window.fig8_female_chart = el;
  }
  "
)

fig8b <- plot_ly(
  df_mye_latest_over_under_65,
  labels = ~age_65split,
  values = ~Males,
  type = "pie",
  text = ~ paste0(
    age_65split,
    " \n (",
    round_half_up(
      Males / sum(Males) * 100,
      1
    ),
    "%)"
  ),
  textinfo = "text",
  hoverinfo = "text",
  # increase label size
  insidetextfont = list(
    size = 16
  ),
  insidetextorientation = "horizontal",
  marker = list(
    colors = c(
      nisra_col3_green,
      nisra_navy
    )
  ),
  # plotly default is grey which is insufficient contrast
  textfont = list(
    color = c(
      "black",
      "white"
    )
  ),
  showlegend = FALSE
) |>
  # modebar is series of icons and plotly logo linked to plotly site. Turned off
  # throughout to avoid multiple links going to same destination (poor
  # accessibility) and icons can create visual clutter.
  config(
    displayModeBar = FALSE
  ) |>
  layout(
    annotations = list(
      x = 0.5,
      y = 1.08,
      xanchor = "center",
      text = "<b> Males </b>",
      showarrow = FALSE,
      font = list(
        family = "Arial",
        size = 16
      )
    ),
    font = list(
      family = "Arial",
      size = 14
    )
  )

fig8b <- htmlwidgets::onRender(
  fig8b,
  "
  function(el, x) {
    window.fig8_male_chart = el;
  }
  "
)

# creates a two column div placing a chart in each. Usual plotly method of
# creating a row with 2 columns did not resize well. This way, the pies display
# one below the other when viewed on mobile.
div(
  class = "row",
  div(
    class = "col-sm-6",
    style = "margin-top: 10px;",
    fig8a
  ),
  div(
    class = "col-sm-6",
    fig8b
  )
)
```

```{r figure_8_dl_buttons}
f_make_tables(
  df_mye_latest_over_under_65,
  title = paste0(
    "Figure 8: Northern Ireland population under 65 and 65 plus by sex ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_ids = c(
    "Female Pie" = "fig8_female_chart",
    "Male Pie" = "fig8_male_chart"
  )
)
```

### Treemap - basic

#### Features

* Added category labels to avoid using a legend.
* Added value labels to help users read the chart.
* Text colours have sufficient contrast against background colours.
* White space between segments acts as a border which means the segments stand
out from one another (this meets minimum contrast).
* Without the white space, the blue and green for example would be harder to
distinguish.

Include descriptive text before or after the chart.

#### Figure 9: Females aged 0-24 make up the largest age group in NI.

##### Female population of Northern Ireland by age group, `r latest_year`

```{r treemap, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 9 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig9 <- plot_ly(
  df_fig9,
  type = "treemap",
  labels = ~age_grp_5,
  parents = ~gender,
  values = ~pop_total,
  # punctuated labels so that readers do not rely solely on colour
  # or shape to understand the chart and screen readers vocalise ok
  text = ~ paste0(
    "Age ",
    age_grp_5,
    ":",
    "\n",
    format(
      pop_total,
      big.mark = ","
    )
  ),
  textinfo = "text",
  hoverinfo = "text",
  hoverlabel = list(
    font = list(
      color = c(
        "white",
        "white",
        "black",
        "white",
        "black"
      )
    )
  ),
  branchvalues = "total",
  # set tile colours, border and label colours to meet accessibility standards
  marker = list(
    colors = c(
      nisra_blue,
      nisra_navy,
      nisra_col3_green,
      nisra_col4_purple,
      nisra_col5_lilac
    ),
    line = list(
      color = "black",
      width = 1
    )
  ),
  textfont = list(
    family = "Arial",
    size = 16,
    color = c(
      "white",
      "white",
      "black",
      "white",
      "black"
    )
  )
)

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig9 <- fig9 |>
  config(
    displayModeBar = FALSE
  )

fig9 <- htmlwidgets::onRender(
  fig9,
  "
  function(el, x) {
    window.fig9_chart = el;
  }
  "
)

fig9
```

```{r figure_9_dl_buttons}
f_make_tables(
  df_fig9_xls,
  title = paste0(
    "Figure 9: Female population of Northern Ireland by age group, ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig9_chart"
)
```
  
### Treemap - nested (3 levels)

This example shows a treemap with lots of boxes and groups. The tree map
(Figure 3) from the Economic and Labour Market Statistics report
[Structure and Performance of the Northern Ireland Economy 2018 and 2019](https://datavis.nisra.gov.uk/economy-and-labour-market/structure-of-the-ni-economy-2019.html)
is shown below. 

#### Figure 10: Split of Northern Ireland GVA by industry, `r df_sut_report_data$Value[1]` 

##### `r highest_gva_industry` accounted for the largest share of GVA

##### <span class="sr-only">Figure 10 tabs</span> {.tabset}

To view the sub-industries in more detail by industry, click on the industry
header at the top of each box. To reset the chart, click on the industry header
again to get back to the original view.

###### `r df_sut_report_data$Value[1]`

```{r figure_10, out.width="100%", echo = TRUE}
tmap_textcolours[1] <- "black"

fig10 <- plot_ly(
  type = "treemap",
  labels = ~tmap_labels_9,
  parents = ~tmap_parents_9,
  values = ~tmap_values_9,
  textinfo = "label",
  text = ~ paste0(
    tmap_labels_9,
    "\n",
    round_half_up(
      tmap_values_9_lab,
      1
    ),
    "%"
  ),
  hoverinfo = "text",

  # ⬇️ put per-tile hover colours HERE (trace-level)
  hoverlabel = list(
    font = list(
      color = tmap_textcolours
    )
  ),
  marker = list(
    colors = c(
      "white",
      nisra_green_decoration,
      nisra_navy,
      nisra_col9_pink,
      nisra_col3_green,
      nisra_col4_purple,
      nisra_col5_lilac,
      nisra_col6_olive,
      nisra_col8_slate,
      nisra_blue,
      nisra_col10_burgandy
    )
  ),
  textfont = list(
    color = tmap_textcolours
  )
) |>
  layout(
    # leave layout.hoverlabel without color (size/bg ok if you want)
    hoverlabel = list(
      font = list(
        size = 20
      )
    ),
    font = list(
      family = "Arial",
      size = 14
    ),
    margin = list(
      l = 0,
      r = 0,
      b = 12,
      t = 0
    )
  ) |>
  config(
    displayModeBar = FALSE
  )

fig10 <- htmlwidgets::onRender(
  fig10,
  "
  function(el, x) {
    window.fig10_chart = el;
  }
  "
)

fig10
```

```{r figure_10_dl_buttons}
f_make_tables(
  data = fig9_dl,
  title = paste0(
    "Figure 10: Split of Northern Ireland GVA by industry, ",
    df_sut_report_data$Value[1]
  ),
  data_style = ns_percentage,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig10_chart"
)
```

###### `r df_sut_report_data$Value[2]`

```{r figure_10_prev, out.width="100%", echo = TRUE}
fig10prev <- plot_ly(
  type = "treemap",
  labels = ~tmap_labels_9prev,
  parents = ~tmap_parents_9prev,
  values = ~tmap_values_9prev,
  textinfo = "label",
  text = ~ paste0(
    tmap_labels_9prev,
    "\n",
    round_half_up(
      tmap_values_9prev_lab,
      1
    ),
    "%"
  ),
  hoverinfo = "text",
  # ⬇️ put per-tile hover colours HERE (trace-level)
  hoverlabel = list(
    font = list(
      color = tmap_textcolours
    )
  ),
  marker = list(
    colors = c(
      "white",
      nisra_green_decoration,
      nisra_navy,
      nisra_col9_pink,
      nisra_col3_green,
      nisra_col4_purple,
      nisra_col5_lilac,
      nisra_col6_olive,
      nisra_col8_slate,
      nisra_blue,
      nisra_col10_burgandy
    )
  ),
  textfont = list(
    color = tmap_textcolours
  )
) |>
  layout(
    hoverlabel = list(
      font = list(
        size = 20
      )
    ),
    font = list(
      family = "Arial",
      size = 14
    ),
    margin = list(
      l = 0,
      r = 0,
      b = 12,
      t = 0
    )
  ) |>
  # modebar is series of icons and plotly logo linked to plotly site. Turned off
  # throughout to avoid multiple links going to same destination (poor
  # accessibility) and icons can create visual clutter.
  config(
    displayModeBar = FALSE
  )

fig10prev <- htmlwidgets::onRender(
  fig10prev,
  "
  function(el, x) {
    window.fig10prev_chart = el;
  }
  "
)

fig10prev
```

```{r figure_10_prev_dl_buttons}
f_make_tables(
  data = fig9prev_dl,
  title = paste0(
    "Figure 10: Split of Northern Ireland GVA by industry, ",
    df_sut_report_data$Value[2]
  ),
  data_style = ns_percentage,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig10prev_chart"
)
```

##### <span class="sr-only">End tabs</span>

### Donut chart

As with pie charts, donut charts should only have 5 categories or fewer. 

#### Features

* The colours are used in the correct order to preserve contrast.
* Segment labels mean there is no need for a legend.
* Text colour of the labels has been set manually to ensure they stand out
(plotly default  does not).
* Labels are short and concise so they remain visible when resized.

#### Figure 11: Under 25s largest age group in Northern Ireland 

##### Northern Ireland population by age group `r latest_year` 

```{r figure_11 donut, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 11 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig11 <- plot_ly(
  df_mye_latest_year_agegroup,
  type = "pie",
  labels = ~age_group,
  values = ~agegroup_total,
  # default orders segments by size
  sort = FALSE,
  # accompanies sort argument
  direction = "clockwise",
  # turns pie into donut
  hole = 0.5,
  # create labels with age group and percentage values in brackets
  text = ~ paste0(
    age_group,
    ": \n (",
    round_half_up(
      agegroup_total / sum(agegroup_total) * 100,
      1
    ),
    "%)"
  ),
  hoverinfo = "text",
  textinfo = "text",
  # keeps labels horizontal
  insidetextorientation = "horizontal",
  # specify colours in correct order.
  marker = list(
    colors = c(
      nisra_blue,
      nisra_navy,
      nisra_col3_green,
      nisra_col4_purple
    ),
    # border required to ensure contrast between nisra blue and lilac
    line = list(
      color = "black",
      width = 1
    )
  ),
  # specify label colours to meet contrast standards
  textfont = list(
    family = "Arial",
    color = c(
      "white",
      "white",
      "black",
      "white"
    ),
    size = 14
  )
) |>
  layout(
    # legend removed for accessibility (relies on colour to be able to
    # understand it. Labels added to segments instead)
    showlegend = FALSE,
    font = list(
      family = "Arial",
      size = 12
    ),
    margin = list(
      r = 70,
      l = 70
    )
  )

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig11 <- fig11 |>
  config(
    displayModeBar = FALSE
  )

fig11 <- htmlwidgets::onRender(
  fig11,
  "
  function(el, x) {
    window.fig11_chart = el;
  }
  "
)

fig11
```

```{r figure_11_dl_buttons}
f_make_tables(
  fig11_xls,
  title = paste0(
    "Figure 11: Northern Ireland population by age group ",
    latest_year
  ),
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig11_chart"
)
```

### Population pyramid 

This population pyramid shows the number of people in each age group. It
compares males and females in two different years that are ten years apart. The
shaded areas show the most recent year. The lines show the earlier year. 

#### Features
* The 4 chart colours have enough contrast to make them easy to see.  
* When you mouse over the chart, the hover label shows all the data for that age
group at the same time. 

Add clear text before or after the chart to explain the main message.

#### Figure 12: The population of Northern Ireland is aging

##### Northern Ireland population by age and sex (`r latest_year - 10` and `r latest_year`)

```{r figure_12 population pyramid, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 12 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig12 <- plot_ly(
  fig_12_data,
  y = ~Age,
  orientation = "h",
  type = "bar",
  showlegend = FALSE,
  hoverinfo = "none"
) |>
  add_trace(
    x = as.formula(
      paste0(
        "~`Males ",
        latest_year,
        "`"
      )
    ),
    name = paste(
      "Males",
      latest_year
    ),
    marker = list(
      color = nisra_navy
    ),
    hoverinfo = "text+y",
    hovertext = paste0(
      "Males ",
      latest_year,
      ": ",
      prettyNum(
        fig_12_data[[
          paste(
            "Males",
            latest_year
          )
        ]],
        big.mark = ","
      )
    )
  ) |>
  add_trace(
    x = as.formula(
      paste0(
        "~`Females ",
        latest_year,
        "`"
      )
    ),
    name = paste(
      "Females",
      latest_year
    ),
    marker = list(
      color = nisra_col3_green
    ),
    hoverinfo = "text+y",
    hovertext = paste0(
      "Females ",
      latest_year,
      ": ",
      prettyNum(
        abs(
          fig_12_data[[
            paste(
              "Females",
              latest_year
            )
          ]]
        ),
        big.mark = ","
      )
    )
  ) |>
  add_trace(
    x = as.formula(
      paste0(
        "~`Males ",
        latest_year - 10,
        "`"
      )
    ),
    type = "scatter",
    mode = "lines",
    name = paste(
      "Males",
      latest_year - 10
    ),
    line = list(
      color = "#00b0f0",
      width = 4
    ),
    showlegend = TRUE,
    hoverinfo = "text+y",
    hovertext = paste0(
      "Males ",
      latest_year - 10,
      ": ",
      prettyNum(
        fig_12_data[[
          paste(
            "Males",
            latest_year - 10
          )
        ]],
        big.mark = ","
      )
    )
  ) |>
  add_trace(
    x = as.formula(
      paste0(
        "~`Females ",
        latest_year - 10,
        "`"
      )
    ),
    type = "scatter",
    mode = "lines",
    name = paste(
      "Females",
      latest_year - 10
    ),
    line = list(
      color = "#2c523b",
      width = 4
    ),
    showlegend = TRUE,
    hoverinfo = "text+y",
    hovertext = paste0(
      "Females ",
      latest_year - 10,
      ": ",
      prettyNum(
        abs(
          fig_12_data[[
            paste(
              "Females",
              latest_year - 10
            )
          ]]
        ),
        big.mark = ","
      )
    )
  ) |>
  layout(
    barmode = "overlay",
    bargap = 0,
    font = list(
      family = "Arial",
      size = 14
    ),
    xaxis = list(
      title = "Persons (Thousands)",
      tickformat = ",",
      tickmode = "array",
      tickvals = seq(
        -15000,
        15000,
        by = 5000
      ),
      ticktext = as.character(
        abs(
          seq(
            -15,
            15,
            by = 5
          )
        )
      ),
      range = c(
        -15000,
        15000
      ),
      showgrid = FALSE,
      zeroline = FALSE, # Remove the internal vertical zero line
      showline = TRUE, # Draw the solid x-axis line
      linecolor = "black",
      linewidth = 1,
      ticks = "outside", # Tick marks on the axis
      ticklen = 8,
      tickwidth = 1,
      tickcolor = "black",
      mirror = FALSE, # Axis line only on the bottom
      anchor = "y", # Tie axis to y-axis (removes gap)
      fixedrange = TRUE,
      domain = c(
        0,
        1
      ), # Ensures the plot spans full width
      constrain = "domain" # Keeps data flush to axes (no extra padding)
    ),
    yaxis = list(
      title = "",
      categoryorder = "category ascending",
      tickvals = seq(
        0,
        90,
        by = 10
      ),
      ticktext = c(
        as.character(
          seq(
            0,
            80,
            by = 10
          )
        ),
        "90+"
      ),
      showgrid = FALSE,
      zeroline = FALSE, # Avoid extra horizontal line
      showline = TRUE, # Draw solid y-axis line
      linecolor = "black",
      linewidth = 1,
      ticks = "outside",
      ticklen = 6,
      tickwidth = 1,
      tickcolor = "black",
      range = c(
        -0.5,
        90.5
      )
    ),
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = 1.1,
      traceorder = "reversed"
    ),
    annotations = list(
      list(
        text = paste0(
          "<b>Females\n",
          latest_year,
          "</b>"
        ),
        x = -5000,
        xref = "x",
        xanchor = "center",
        y = 30,
        yref = "y",
        yanchor = "middle",
        showarrow = FALSE,
        font = list(
          color = "#000",
          size = 16
        )
      ),
      list(
        text = paste0(
          "<b>Males\n",
          latest_year,
          "</b>"
        ),
        x = 5000,
        xref = "x",
        xanchor = "center",
        y = 30,
        yref = "y",
        yanchor = "middle",
        showarrow = FALSE,
        font = list(
          color = "#fff",
          size = 16
        )
      )
    ),
    hovermode = "y unified"
  )

# modebar is series of icons and plotly logo linked to plotly site. Turned off
# throughout to avoid multiple links going to same destination (poor
# accessibility) and icons can create visual clutter.
fig12 <- fig12 |>
  config(
    displayModeBar = FALSE
  )

fig12 <- htmlwidgets::onRender(
  fig12,
  "
  function(el, x) {
    window.fig12_chart = el;
  }
  "
)

fig12
```

```{r figure_12_dl_buttons}
f_make_tables(
  fig_12_xl,
  title = paste0(
    "Figure 12: Population by age and sex (",
    latest_year - 10,
    " and ",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig12_chart"
)
```

### Flow chart

This flow chart shows how the population in Northern Ireland changed between
2021 and 2022. The chart was created in the Markdown document using shapes
defined in the CSS file.

#### Features
* Background colours are all light to ensure black text stands out and is easy
to read.
* If you change the colours, make sure the text and background still have enough
contrast to meet accessibility standards.

Add clear text before or after the chart to explain the main message.

#### Figure 13: Natural change is the main driver of population growth

##### Components of population change in Northern Ireland (`r prev_year` and `r latest_year`)

```{r fig13, echo=TRUE, results='asis'}
cat(
  sprintf(
    '
::::::::: div-row
::: {.navy-box style="font-weight: bold"}
%s\\
Population\\
(%s)
:::

::: arrow

:::

::: green-circle
Natural\\
Change\\
%s
:::

::: blue-circle
Net\\
Migration\\
%s
:::

::: green-circle
Other\\
Changes\\
%s
:::

::: arrow

:::

::: {.navy-box style="font-weight: bold"}
%s\\
Population\\
(%s)
:::
:::::::::
',
    prev_year,
    pop_ni_last_rounded,
    natural_change_sign,
    net_migration_sign,
    other_changes_sign,
    latest_year,
    pop_ni_current_rounded
  )
)
```

### Filled multiple line chart

This chart shows population change in Northern Ireland over the course of five
years by five broad age bands. Labels are placed at the start and end of each
time series to avoid visual clutter.

#### Figure 14: Over 65s see largest population growth

##### Northern Ireland population change by age group (`r latest_year - 4` to `r latest_year`)

```{r fig14 filled multiple line chart, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Figure 14 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig_14_plot <- plot_ly(
  data = fig_14_data,
  x = ~index,
  y = ~pct,
  fill = "tozeroy",
  type = "scatter",
  mode = "none",
  color = ~age_grp_5,
  fillcolor = nisra_blue,
  hoverinfo = "text",
  text = ~ paste(
    mid_year_ending,
    age_grp_5,
    paste0(
      round_half_up(
        pct,
        1
      ),
      "%"
    ),
    sep = "\n"
  )
) |>
  layout(
    font = list(
      family = "Arial",
      size = 12
    ),
    title = "",
    xaxis = list(
      title = "",
      showticklabels = FALSE,
      showgrid = FALSE
    ),
    yaxis = list(
      title = "",
      range = c(
        0,
        35
      ),
      showgrid = FALSE, # No vertical gridlines
      zeroline = TRUE, # Keep the main x-axis line
      showline = TRUE, # Ensure axis line is drawn
      linecolor = "black", # Axis line color
      linewidth = 1
    ),
    showlegend = FALSE,
    margin = list(
      b = 100
    ),
    hovermode = "x unified"
  )

fig_14_data$age_grp_5 <- as.factor(
  fig_14_data$age_grp_5
)

age_groups <- c(
  "Under 24",
  "25 to 44",
  "45 to 64",
  "65 to 84",
  "85 and over"
)

x_positions <- seq(
  1 / (length(age_groups) * 2),
  (length(age_groups) * 2 - 1) /
    (length(age_groups) * 2),
  length.out = length(age_groups)
)

annotations <- lapply(
  seq_along(age_groups),
  function(i) {
    list(
      text = age_groups[i],
      x = x_positions[i],
      xref = "paper",
      xanchor = "center",
      y = -0.1,
      yref = "paper",
      yanchor = "top",
      showarrow = FALSE
    )
  }
)

annotations_left <- list()
annotations_right <- list()

for (i in seq_len(nrow(fig_14_data))) {
  if (fig_14_data$index[i] %% 5 == 1) {
    annotations_left[[i]] <- list(
      x = fig_14_data$index[i],
      y = 0,
      text = fig_14_data$label[i],
      xanchor = "left",
      yanchor = "top",
      xref = "x",
      yref = "paper",
      showarrow = FALSE
    )
  } else if (fig_14_data$index[i] %% 5 == 0) {
    annotations_right[[i]] <- list(
      x = fig_14_data$index[i],
      y = 0,
      text = fig_14_data$label[i],
      xanchor = "right",
      yanchor = "top",
      xref = "x",
      yref = "paper",
      showarrow = FALSE
    )
  }
}

fig_14_plot <- fig_14_plot |>
  layout(
    annotations = c(
      annotations,
      annotations_left,
      annotations_right
    )
  )

for (i in seq_len(nrow(fig_14_data))) {
  if (fig_14_data$mid_year_ending[i] == year_minus_four) {
    fig_14_plot <- fig_14_plot |>
      add_annotations(
        text = paste0(
          sprintf(
            "%0.1f",
            fig_14_data$pct[i]
          ),
          "%"
        ),
        x = fig_14_data$index[i],
        y = fig_14_data$pct[i] + 1.4,
        xanchor = "left",
        yanchor = "bottom",
        showarrow = FALSE
      )
  }

  if (fig_14_data$mid_year_ending[i] == latest_year) {
    fig_14_plot <- fig_14_plot |>
      add_annotations(
        text = paste0(
          sprintf(
            "%0.1f",
            fig_14_data$pct[i]
          ),
          "%"
        ),
        x = fig_14_data$index[i],
        y = fig_14_data$pct[i] + 1.4,
        xanchor = "right",
        yanchor = "bottom",
        showarrow = FALSE
      )
  }
}

fig_14_plot <- fig_14_plot |>
  f_responsive_annotations()

fig14 <- fig_14_plot |>
  config(
    displayModeBar = FALSE
  )

fig14 <- htmlwidgets::onRender(
  fig14,
  "
  function(el, x) {
    window.fig14_chart = el;
  }
  "
)

fig14
```

```{r figure_14_dl_buttons}
f_make_tables(
  fig_14_data_xl,
  title = paste0(
    "Figure 14: Northern Ireland population change by age group (",
    latest_year - 4,
    " to ",
    latest_year,
    ")"
  ),
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig14_chart"
)
```

###  Diverging stacked bar 

This chart has been included to show how to display data with positive and
negative change.

#### Figure 15: Hawaii had the largest positive population change

##### Population change in selected US States in 2022

```{r figure_15, echo = TRUE, fig.alt="Figure 15 alt text example. More detail on the trends are included in the text directly above or below the chart, and data can be downloaded using the button below", out.width="100%"}
fig15 <- plot_ly(
  fig_15_data_filtered,
  x = ~`Net Migration & Other Changes`,
  y = ~`Local Government District`,
  type = "bar",
  orientation = "h",
  marker = list(
    color = nisra_col3_green
  ),
  name = "Net migration and other changes",
  hoverinfo = "text+x",
  hovertext = ~ paste(
    "Net migration:",
    `Net Migration & Other Changes`
  )
) |>
  add_trace(
    x = ~`Natural Change (births minus deaths)`,
    y = ~`Local Government District`,
    type = "bar",
    orientation = "h",
    marker = list(
      color = nisra_navy
    ),
    name = "Natural change (births minus deaths)",
    hovertext = ~ paste(
      "Natural change:",
      `Natural Change (births minus deaths)`
    )
  ) |>
  add_trace(
    x = ~`Total Percentage Change`,
    y = ~`Local Government District`,
    type = "scatter",
    mode = "markers",
    marker = list(
      color = nisra_blue,
      size = 9,
      symbol = "diamond",
      line = list(
        color = "#fff",
        width = 2
      )
    ),
    name = "Total percentage change",
    hovertext = ~ paste(
      "Total change:",
      `Total Percentage Change`
    )
  ) |>
  layout(
    font = list(
      family = "Arial",
      size = 14
    ),
    barmode = "relative",
    xaxis = list(
      title = "Percentage change",
      range = c(
        -50,
        50
      ),
      tickmode = "array", # manually define ticks
      tickvals = c(
        -50,
        -25,
        0,
        25,
        50
      ), # exact tick positions
      ticktext = c(
        "-50%",
        "-25%",
        "0%",
        "25%",
        "50%"
      ), # custom labels (optional)
      automargin = TRUE,
      tickfont = list(
        size = 14
      )
    ),
    yaxis = list(
      title = "",
      automargin = TRUE
    ),
    legend = list(
      orientation = "h",
      y = 1.35,
      xanchor = "center"
    ),
    hovermode = "y unified",
    margin = list(
      l = 10,
      r = 10,
      t = 10,
      b = 60
    )
  )

fig15 <- fig15 |>
  config(
    displayModeBar = FALSE
  )

fig15 <- htmlwidgets::onRender(
  fig15,
  "
  function(el, x) {
    window.fig15_chart = el;
  }
  "
)

fig15
```

```{r figure_15_dl_buttons}
f_make_tables(
  fig_15_data_filtered,
  title = "Figure 15: Population change in selected US States in 2022",
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig15_chart"
)
```

### Other chart types

There are many other charts and maps to choose from that can be coded in R and
included in html reports. More information and examples can be explored in the
following links:

[Plotly R Open Source Graphing Library](https://plotly.com/r/#:~:text=Plotly's%20R%20graphing%20library%20makes,3D%20(WebGL%20based)%20charts.)

[Plotly R Library Basic Charts](https://plotly.com/r/basic-charts/)

[Plotly R Library Statistical Charts](https://plotly.com/r/statistical-charts/)

[The R Graph Gallery](https://r-graph-gallery.com/index.html)

[The R Graph Gallery - Interactive Charts](https://r-graph-gallery.com/interactive-charts.html)

[Plotly cheat sheet for R (PDF, 2.6 MB)](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)

[ggplot2 charts](https://r-charts.com/ggplot2/)

[Data visualization with ggplot2](https://rstudio.github.io/cheatsheets/html/data-visualization.html)

[Making Maps with R](https://www.geeksforgeeks.org/making-maps-with-r/)

[Using ggplot2 to create maps](https://bookdown.org/nicohahn/making_maps_with_r5/docs/ggplot2.html)

[Sankey Diagram](https://r-graph-gallery.com/sankey-diagram.html)

[Building a flowchart](https://cran.r-project.org/web/packages/Gmisc/vignettes/Grid-based_flowcharts.html)
