---
title: "06 Map Sub Report"
output: html_document
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_dir = "../demo/demo_outputs/",
    output_file = paste0(
      xfun::sans_ext(input),
      "_",
      format(Sys.time(), "%d-%m-%Y_%H%M"),
      ".html"
    ),
    envir = globalenv()
  )
})
---

```{r maps_settings, include=FALSE}
# Set chunk options for all chunks in this file.

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)

library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

This section includes a static map and an interactive map. If you create a map
in another package and import as an image, ensure it meets accessibility
requirements for an infographic.

### Static map 

This section includes a map generated in R.

#### Accessible maps

*	Make sure different colours are easy to tell apart.
* More colours can make the map harder to read.
* Make sure the contrast between text labels and the background colour is at
least 4.5:1.
* Try to make sure colours next to each other have a colour contrast of at least
3:1.
* If shape colours do not have enough contrast, use borders that do.
* Check colours using tools such as the
[WebAIM contrast checker](https://webaim.org/resources/contrastchecker/)
* Do not rely on colour alone to communicate information (e.g. legends).
* Use clear and descriptive labels. Do not rely on hover interactions alone to
communicate important information.

#### Features of this static map

* Labels are shown on the map.
* Alt text has been added for screen reader users.
* The shapefile has been made smaller so it loads more quickly.
* Only three colours are used.
* There is no map background so the shapes are more distinct.
* Data is available to download for users who cannot see the map.


Add a short description before or after the map to explain what it shows.

#### Map 1: Proportion of people aged under 25 in Northern Ireland by LGD

```{r map_static, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Map of Northern Ireland shaded by percentage of under 25s living in each Local Government District. More detail on the trends are included in the text directly above or below the map", out.width="100%"}
df_map_data$xmod <- c(
  0.8,
  0,
  0,
  0.5,
  0,
  1.1,
  0.5,
  0.5,
  0,
  1,
  -1
)

df_map_data$ymod <- c(
  -0.5,
  0,
  0.2,
  0.3,
  0,
  0,
  1.4,
  -0.55,
  0.3,
  0,
  -1
)

# Plotting map:
tmap_mode("plot") # Set to "view" for interactive map or "plot" for static

map1_static_map <- tm_shape(
  df_map_data,
  # Default is km
  unit = "imperial"
) +
  tm_polygons(
    # Colours the map sections based on these values
    col = "lgd_young_perc",
    # Legend title
    title = "Percentage of under 25s",
    # Number of different colours to use
    n = 3,
    # Quantiles gives even sized groupings
    style = "quantile",
    # Colour palette. Execute the command tmaptools::palette_explorer() in R
    # console to see range of options
    palette = "Blues",
    # Add transparency to polygons improving contrast with text
    alpha = 0.5,
    legend.show = FALSE,
    border.col = "black"
  ) +
  tm_text(
    # add labels to polygons, change text colour to black and move
    # labels to fit within shapes
    "map_label",
    size = 0.8,
    col = "#000000",
    auto.placement = FALSE,
    xmod = "xmod",
    ymod = "ymod"
  ) +
  tm_add_legend(
    type = "polygons",
    fill = RColorBrewer::brewer.pal(
      3,
      "Blues"
    ),
    col = "#454A58",
    labels = c(
      "26.6 to 29.5",
      "29.6 to 31.5",
      "31.6 to 33.5"
    ),
    title = "Percentage of under 25s"
  ) +
  tm_add_legend(
    # add manual legend as a key
    "symbol",
    fill = "#00205b",
    title = "Key",
    labels = c(
      "AN = Antrim and Newtownabbey",
      "AND = Ards and North Down",
      "B = Belfast"
    )
  ) +
  tm_format(
    # setting format to worldwide creates a wider rectangle shape that
    # fits width of page
    "World_wide"
  ) +
  tm_layout(
    legend.position = c(
      "left",
      "top"
    )
  )

map1_static_map

tmap_save(
  tm = map1_static_map,
  filename = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata",
    "map1_static.png"
  ),
  width = 8,
  height = 6,
  units = "in"
)

map_download_data <- df_map_data |>
  sf::st_drop_geometry() |>
  select(
    lgd2014name,
    lgd_young_perc
  )

f_make_tables(
  data = map_download_data,
  title = "Map 1",
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  ggplot = TRUE
)
```

### Interactive maps

This section shows the same map produced as an interactive plot.

#### Features of interactive map

* The values are displayed on the map so that the user is not relying on a
legend.
* The background map is greyscale and the detail is limited to help the shapes
stand out.
* Council names and figures display on click.
* There was insufficient space to show council names on the map.
* The shapefile has been simplified to reduce file size and not slow page load
speed.
* Data is available to download for users who cannot see the map.
* The default rendering of the map tabbed through each LGD shape, but without
any highlighting or visible indication of this so it failed accessibility
testing.
* Code has been added to remove the map shapes from the tab order to address
this accessibility failure (see end of demo_report.Rmd)


Include descriptive text before or after the map.

#### Map 2: Interactive LGD map with additional information

The map shows the percentage of people aged under 25 in each Northern Ireland
Local Government District. Values are also displayed directly on the map and are
available in downloadable CSV and Excel formats below.

```{r map_interactive_1, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Map of Northern Ireland shaded by percentage of under 25s living in each Local Government District. More detail on the trends are included in the text directly above or below the map", out.width="100%"}
df_map_data$xmod <- c(
  0,
  -1.5,
  0,
  -1.1,
  0,
  0,
  0,
  -1,
  0,
  0,
  -1
)

df_map_data$ymod <- c(
  0,
  1,
  0,
  0,
  0,
  0,
  1,
  0,
  0,
  0,
  -1
)

tmap_mode(
  # Set to "view" for interactive map or "plot" for static
  "view"
)

map2_interactive_map <- tm_shape(
  df_map_data |>
    select(
      lgd2014name,
      lgd_young_perc,
      map_label_value_only,
      ymod,
      xmod
    ),
  name = "Show data",
  # Default is km
  unit = "imperial"
) +
  tm_basemap(
    # Adds background map.
    "Esri.WorldGrayCanvas"
  ) +
  tm_polygons(
    # Colours the map sections based on these values
    col = "lgd_young_perc",
    # Number of different colours to use
    n = 3,
    # Quantiles gives even sized groupings
    style = "quantile",
    # Colour palette
    palette = "Blues",
    # Add transparency to polygons improving contrast with text
    alpha = 0.5,
    border.alpha = 1,
    legend.show = FALSE,
    popup.vars = c(
      "Council" = "lgd2014name",
      "Percentage under 25" = "lgd_young_perc"
    )
  ) +
  tm_add_legend(
    type = "polygons",
    fill = RColorBrewer::brewer.pal(
      3,
      "Blues"
    ),
    col = "#454A58",
    lwd = 1,
    labels = c(
      "26.6 to 29.5",
      "29.6 to 31.5",
      "31.6 to 33.5"
    ),
    title = "Percentage of under 25s"
  ) +
  tm_view(
    # Re-positioning legend
    view.legend.position = c(
      "left",
      "top"
    ),
    # Re-positioning layer controls
    control.position = c(
      "left",
      "bottom"
    ),
    leaflet.options = c(
      zoomControl = FALSE,
      # Removes zoom and drag options
      dragging = FALSE,
      keyboard = FALSE,
      boxZoom = FALSE,
      doubleClickZoom = FALSE,
      scrollWheelZoom = FALSE,
      tap = FALSE,
      touchZoom = FALSE,
      zoom = 8,
      minZoom = 8,
      maxZoom = 8,
      attributionControl = FALSE
    )
  ) +
  tm_text(
    # Label polygons, increase font size, set colour to black (better
    # contrast) and adjust positioning so they are contained as much as
    # possible within the polygons
    "map_label_value_only",
    size = 1.2,
    col = "#000000",
    auto.placement = FALSE,
    xmod = "xmod",
    ymod = "ymod"
  ) +
  tm_scalebar(
    width = 250
  )

map2_interactive_map

map_download_data <- df_map_data |>
  sf::st_drop_geometry() |>
  select(
    lgd2014name,
    lgd_young_perc
  )

f_make_tables(
  data = map_download_data,
  title = "Map 2 Data",
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  )
)

# please also refer to the "alt text for maps" chunk in the source code of this
# section for additional configurations
```

#### Interactive point map with zoom and hover text

Map 3 shows an interactive map with points marked on it instead of areas. The
map uses data about Jobs and Benefits Offices (JBOs) and the number of job
vacancies at each office.

#### Features

* Users can hover over or select a point to see the office name and the number
of vacancies.
* Users can zoom in and out and reset the view.
* The points can be made larger or smaller based on the data, or grouped
together when many points are close to each other.
* The data is available to download or view as a table on the next tab.

The interactive map is not fully accessible to keyboard users. The underlying
data are provided in the adjacent data tab and as downloadable files, although
these alternatives do not communicate the spatial relationships shown by the
map.

#### Map 3: Vacancies at Jobs and Benefits Offices

##### <span class="sr-only">Map 3 tabs</span> {.tabset}

###### Map (interactive) output

```{r map_interactive_2, echo = TRUE, warnings = FALSE, message = FALSE, fig.alt="Map of Northern Ireland with points showing the location of benefits offices (JBO) and the number of available vacancies. More detail on the trends are included in the text directly above or below the map", out.width="100%"}
#### interactive map ####

map3_interactive_map <- leaflet(
  options = leafletOptions(
    minZoom = 8,
    maxZoom = 12
  )
) |>
  setView(
    lng = -6.5,
    lat = 54.6,
    zoom = 8.15
  ) |>
  addProviderTiles(
    "Esri.WorldGrayCanvas"
  )

map3_interactive_map <- addCircleMarkers(
  map = map3_interactive_map,
  data = df_jbo,
  lng = ~Longitude,
  lat = ~Latitude,
  color = "blue",
  radius = 10,
  label = lapply(
    paste0(
      df_jbo$JBO,
      ": ",
      df_jbo$Vacancies,
      " vacancies"
    ),
    HTML
  ),
  labelOptions = labelOptions(
    style = list(
      "font-size" = "14px"
    )
  )
) |>
  addEasyButton(
    easyButton(
      icon = "fa-globe",
      title = "Reset Zoom",
      onClick = JS("function(btn, map) { map.setView([54.6, -6.5], 8.15); }")
    )
  )

map3_interactive_map
```

```{r map3_dl_buttons}
map3_download_data <- df_jbo |>
  select(
    JBO,
    Vacancies
  )

f_make_tables(
  data = map3_download_data,
  title = "Map 3 Data",
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  )
)
```

###### Map (interactive) data

```{r map_interactive_2b, echo = TRUE}
kable(
  map3_download_data,
  col.names = c(
    "Jobs and Benefits Office",
    "Vacancies"
  )
)
```

```{r alt_text_for_maps, echo = FALSE}
# Alt text values for tabRemove() function in the interactive map
map_alt <- paste0(
  df_map_data$lgd2014name,
  " ",
  df_map_data$map_label_value_only
)

map_alt <- paste(
  map_alt,
  collapse = '", "'
)

map_alt <- paste0(
  '["',
  map_alt,
  '"]'
)
```  

##### <span class="sr-only">End tabs</span>
