---
title: "07 Download Data Sub Report"
output: html_document
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_dir = "../demo/demo_outputs/",
    output_file = paste0(
      xfun::sans_ext(input),
      "_",
      format(Sys.time(), "%d-%m-%Y_%H%M"),
      ".html"
    ),
    envir = globalenv()
  )
})
---

```{r download_data_settings, include=FALSE}
# Set chunk options for all chunks in this file.

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

### Download buttons linked to charts and maps

The make_tables.R script in the functions folder contains the f_make_tables()
each figure and map in Sections 5 and 6. The drop-down provides options to
download the data as an Excel or CSV file, or to download an image of the chart
or map as a PNG file.

The Excel and CSV files are created and saved to the figdata folder when the
report is generated. When a user selects the PNG download option, the image is
generated and saved to the user's Downloads folder.

The f_make_tables function can be added after each figure or map. The main
inputs used in this example are:

* data - the data used in the chart or map.
* title - the title used to create the downloaded filenames.
* data_style - the style applied to the data in the Excel file. Available
styles can be found in the Style.R script.
* data_dir - the folder where the Excel and CSV files are saved. In this
example, the file path is set in the demo configuration file.
* plotly_id - the ID used to identify the Plotly chart or map when generating
the PNG image.

An example of how f_make_tables is used for Figure 1 is shown below. Select
**Show code** to view the example.
```{r data_download_example, echo = TRUE}
# call make tables function to create chart download buttons
f_make_tables(
  data = df_fig1_xls,
  title = fig1_title,
  data_style = ns_comma,
  data_dir = here(
    "code",
    "demo",
    "demo_outputs",
    "figdata"
  ),
  plotly_id = "fig1_chart"
)
```

### Embed spreadsheet into report 

Another way to provide downloadable data in an HTML report is to add a "Download
data" link that links to an embedded spreadsheet.

To use this option, first run the demo_excel_tables.R script to create the
spreadsheet in the demo output folder. You can also add an existing spreadsheet
to the folder.

The embed_file function is then used to create the download button. You can set:

*	the file path
*	the file name
*	the text shown on the button

Links to downloads should have file type and size in the link text [2022 Mid-year population estimates statistical bulletin pdf 545KB](https://www.nisra.gov.uk/system/files/statistics/Statistical%20Bulletin%20-%202022%20Mid-year%20Population%20Estimates%20for%20Northern%20Ireland.pdf).

Select Show Code below to see an example.

This option should only be used when needed. Embedded spreadsheets can make the
HTML file much larger. Large files can take longer to process, create and load.
They can also make it harder to collect usage statistics for the downloaded
files. 

Where possible, store spreadsheets outside the HTML report and add a link
to them instead. This keeps the report smaller, faster to load and allows
download statistics to be collected.


```{r data_download, echo =TRUE, eval=demo_excel_ready}
# embeds the spreadsheets
embexcel <- embed_file(
  here(
    "code",
    "demo",
    "demo_outputs",
    "RAP_demo_tables.xlsx"
  ),
  text = "Publication tables: 15KB"
)

div(
  div(
    class = "row",
    style = "display: flex;",
    div(
      class = "row-indent"
    ),
    div(
      class = "download",
      "Download data: ",
      style = "font-size: 12pt"
    ),
    div(
      class = "row-indent"
    ),
    div(
      class = "xl-button2",
      HTML(embexcel) # <-- wrap with HTML()
    ),
    div(
      class = "download",
      style = paste0(
        "padding-top: 7px; ",
        "padding-left: 5px; ",
        "padding-right: 5px;"
      )
    )
  )
)
```
