---
title: "09 Text Presentation Sub Report"
output: html_document
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_dir = "../demo/demo_outputs/",
    output_file = paste0(
      xfun::sans_ext(input),
      "_",
      format(Sys.time(), "%d-%m-%Y_%H%M"),
      ".html"
    ),
    envir = globalenv()
  )
})
---

```{r text_presentation_settings, include=FALSE}
# Set chunk options for all chunks in this file.

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

This section includes a quote, more subheading levels, unordered lists, ordered
lists and links.

### Quote box

<!-- coloured definiton boxes -->
<p class="defboxes">
  <font color = `r nisra_navy`><strong>Mid Year Estimates:</strong>    
    Every June, NISRA counts how many people live in Northern Ireland. It also
    shares information about different places and areas across the country. A
    style like this can be used to explain important words or to make something
    stand out.
  </font>
</p> 

### Areas of Northern Ireland 

#### Local government districts 

  1. Antrim and Newtownabbey (no ampersand).
  2. Ards and North Down.
  3. Armagh City, Banbridge and Craigavon.
  4. Belfast.
  5. Causeway Coast and Glens.
  6. Derry City and Strabane.
  7. Fermanagh and Omagh.
  8. Lisburn and Castlereagh.
  9. Mid and East Antrim.
  10. Mid Ulster.
  11. Newry, Mourne and Down.

#### Health and social care trusts in Northern Ireland

  1. Belfast HSC Trust.
  2. South Eastern HSC Trust.
  3. Western HSC Trust.
  4. Southern HSC Trust.
  5. Northern HSC Trust.
  6. Northern Ireland Ambulance Service.

#### Geography levels in Northern Ireland 

  * NUTS III areas - large parts of Northern Ireland that are used to group
  places together and compare information.
  * Super Output Areas - small local areas that help us understand what is
  happening in different neighbourhoods.
  * Census Small Areas - very small areas made up of a few streets or homes.
  They help us to count people and learn about local communities.

#### Former areas and health trusts in Northern Ireland 

  -	Parliamentary Constituencies (Assembly Areas). 
  -	Health and Social Care Trusts. 
  -	The former Education and Library Boards. 
  -	The former Health and Social Services Boards.
  - The former NUTS-III areas. 
  -	The former 582 Electoral Wards.


### Using bullet points.
There are two ways to use bullet points.

#### After a heading

- Start each bullet with a capital letter.
- End each bullet with a full stop.
-	Keep each bullet short.

In a list within a sentence:

- add a short lead-in line before the list
- start each bullet with a lowercase letter, unless it starts with a name
-	do not use full stops at the end of the bullets

For more help, see the ONS punctuation guide.

### Accordions

Accordions are useful when you want to switch between hiding and showing a large
amount of content.

Notes can be placed inside an accordion, like the example below. Accordions help
keep pages tidy. Users can open them when they want to read more information.


```{r results='asis', echo = TRUE}
htmltools::HTML(
  "
  <div class='accordion'>
    <button class='accordion-toggle' aria-expanded='false'
            aria-controls='panel1' id='accordion1'>
      Local Government Districts
      <span class='accordion-arrow'>⌵</span>
    </button>
    <div class='accordion-panel' id='panel1' role='region'
         aria-labelledby='accordion1'>

      <ul>
        <li>Antrim and Newtownabbey (no ampersand).</li>
        <li>Ards and North Down.</li>
        <li>Armagh City, Banbridge and Craigavon.</li>
        <li>Belfast.</li>
        <li>Causeway Coast and Glens.</li>
        <li>Derry and Strabane.</li>
        <li>Fermanagh and Omagh.</li>
        <li>Lisburn and Castlereagh.</li>
        <li>Mid and East Antrim.</li>
        <li>Mid Ulster.</li>
        <li>Newry, Mourne and Down.</li>
      </ul>

    </div>
  </div>

  <script>
  document
    .getElementById('accordion1')
    .addEventListener('click', function() {
      const expanded =
        this.getAttribute('aria-expanded') === 'true';
      const panel =
        document.getElementById(this.getAttribute('aria-controls'));

      this.setAttribute('aria-expanded', !expanded);

      if (!expanded) {
        panel.style.maxHeight = panel.scrollHeight + 'px';
        panel.style.opacity = '1';
        panel.style.transform = 'translateY(0)';
      } else {
        panel.style.maxHeight = '0';
        panel.style.opacity = '0';
        panel.style.transform = 'translateY(-10px)';
      }
    });
  </script>
  "
)
```

### Hyperlinks

A hyperlink is a piece of text which allows people to navigate between different
web pages. When adding a link, do not use the web address as the link text.

Use words that tell users where the link will take them. If the link is to a
file, include the file type and size in the link text.

For data downloads, it is usually better to link to the publication page rather
than the file itself.

If you link directly to a file:

- include the file type and size in the link text
- tell users if the file opens in a new window

The hyperlink colour in this demo report was changed because the default colour
did not have enough contrast to meet accessibility standards.

#### Examples of adding links 

For a link that opens in the same browser window, put the link name in square
brackets. Then put the url in round brackets immediately after. For
accessibility links should open in the same window.  

[NISRA - 2022 Mid Year Population Estimates for Northern Ireland](https://www.nisra.gov.uk/publications/2022-mid-year-population-estimates-northern-ireland)  

```markdown
[NISRA - 2022 Mid Year Population Estimates for Northern Ireland](https://www.nisra.gov.uk/publications/2022-mid-year-population-estimates-northern-ireland)
```
Where a link will open in a new window the "(opens in a new window)" text should
be added to the end of the link name. In addition an icon to show the user it
opens in a new window can be added with R markdown code such as
`<span class="new-tab-icon"></span>)`. When target = "_blank" is set the link
opens in a new window or tab.  

[NISRA - 2022 Mid Year Population Estimates for Northern Ireland (opens in a new window) <span class="new-tab-icon"></span>](https://www.nisra.gov.uk/publications/2022-mid-year-population-estimates-northern-ireland){target="_blank"}  

```markdown
[NISRA - 2022 Mid Year Population Estimates for Northern Ireland (opens in a new window) <span class="new-tab-icon"></span>](https://www.nisra.gov.uk/publications/2022-mid-year-population-estimates-northern-ireland){target="_blank"}
```
  
If links take you directly to a download or file, then the link text should have
the file type and size included below.  

[2022 Mid-year population estimates statistical bulletin pdf 545KB](https://www.nisra.gov.uk/system/files/statistics/Statistical%20Bulletin%20-%202022%20Mid-year%20Population%20Estimates%20for%20Northern%20Ireland.pdf)  

```markdown
[2022 Mid-year population estimates statistical bulletin pdf 545KB](https://www.nisra.gov.uk/system/files/statistics/Statistical%20Bulletin%20-%202022%20Mid-year%20Population%20Estimates%20for%20Northern%20Ireland.pdf)
```
