---
title: "NISRA Technology and Support Lab RAP Skeleton Demo Quality Assurance (QA) Report"
lang: "en"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_download: no
    css: "../style.css"
    includes:
      in_header:
        - "../consent_head.html"
      after_body:
        - "../cookie_banner.html"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../demo/demo_outputs/",
      output_file = paste0(
        xfun::sans_ext(input),
        "_",
        format(
          Sys.time(),
          "%d-%m-%Y_%H%M"
        ),
        ".html"
      ),
      envir = globalenv()
    )
  })
params:
  annotations: true
---

<div id="cookie-banner"></div>

```{r settings, include=FALSE}
# this chunk sets chunk options for all chunks in this file
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)
```

```{r setup, warnings = FALSE}
library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)
```

<div data-nics-theme="`r nics_theme`">

<!-- open main div -->

<main id="main">

## Overall Figures

**This Quality Assurance (QA) report uses the mid-year population estimates**
**data used in the `demo_report.Rmd` as an exemplar and provides guidance for**
**what to put in each section of the QA report**.

Use the "Overall Figures" section of the QA report to point out the key figures
throughout the report. This should include the total number of records. If
needed, also show how the numbers are split by time period, such as year,
quarter, or month. See the examples below:

- Total number of records in the dataset is
`r format(total_records_myes, big.mark=",")`.
- Total population count in `r earliest_year` across all councils is
`r format(population_est_2001, big.mark=",")`.
- Total population count in `r latest_year` across all councils is
`r format(population_est_2022, big.mark=",")`.
- Population difference between `r latest_year` and `r earliest_year` across all
councils is
`r format(abs(population_est_2022 - population_est_2001), big.mark=",")`.

## Compare Trends

This section of the QA report should show the main trends in the data. This may
include population figures from the current year and previous year. Highlight
the key numbers and explain how they have changed over time. Use h3 (heading
level 3) to label each table. Examples of population changes for males and
females are shown below:

### Male population breakdown by LGDs

Table showing the change in the population of males in the top 3 LGDs for the
current and previous year. Use  h3 for labelling the heading level of the tables.

``` {r, out.width="100%"}
kable(
  prev_curr_lgd_breakdown_male,
  align = ("lrrrr")
)
```

### Female population breakdown by LGDs

Table showing the change in the population of females in the top 3 LGDs for the
current and previous year.

``` {r, out.width="100%"}
kable(prev_curr_lgd_breakdown_female,
  align = ("lrrrr")
)
```

## Counts by variables 

This section of the QA report is for reporting on the counts of key variables
in the data. The report uses mid-year population estimates and the population
breakdowns for some of the variables present in the data.

### Male and Female population

Table showing the variable count of the population of males and females in
`r latest_year`.

``` {r, out.width="100%"}
kable(
  df_latest_year_gender_cleaned,
  align = ("lrrr")
)
```

### Population breakdown for NI population by LGDs in `r latest_year`

Table showing the population count per LGD for all persons of the NI population
in `r latest_year`.

``` {r, out.width="100%"}
kable(
  df_mye_latest_ldg_cleaned,
  align = ("lrrr")
)
```

### Population breakdown for NI population by LGD in `r earliest_year`

Table showing the population count breakdown by LGD for all persons of the NI
population in `r earliest_year`

``` {r, out.width="100%"}
kable(
  df_mye_earliest_ldg,
  align = ("lrrr")
)
```

### Population differences for NI population by LGD in `r latest_year - 1` and `r latest_year`

Table showing the population breakdown by LGD for all persons of the NI
population. This table also flags any unusual percentage changes (above or below
0.5%) between the population in consecutive years to identify outliers in the
data.

``` {r, out.width="100%"}
kable(
  df_mye_curr_prev_population,
  align = ("lrrrr")
)
```

### Population breakdown for NI population by age in `r latest_year`

Table showing the age breakdown for all persons of the NI population in
`r latest_year`.

``` {r, out.width="100%"}
kable(
  df_latest_year_age_group_clean,
  align = ("lrrr")
)
```

### Population breakdown for NI population by age in `r earliest_year`

Table showing the age breakdown for all persons of the NI population in
`r earliest_year`.

``` {r, out.width="100%"}
kable(
  df_mye_earliest_year_agegroup,
  align = ("lrrr")
)
```

## Suppression

Suppression is used to protect confidentiality by masking small values that
could potentially identify individuals. This section reports the outcome of
checks applied to publication tables following suppression.

List each table in the report using a h3 heading. Check whether suppression has
been applied and record the result under the table heading. Also include the
name of the variable used to create the table. Use the example below as a guide:

### Table 1a: population of young people by LGD

`df_mye_latest_year_youthrate` (the name of the dataframe) has no NAs.

### Table 1b: population of elderly people by LGD

`df_mye_latest_year_elderlyrate` has the suppression function applied.

Number of suppressed cells: 3.

### Table 2

`df_mye_year_gender_t` has no NAs or suppression applied.

<!-- close main div -->

</main>

```{r render-footer, results='asis', echo=FALSE}
f_footer()
```

`r if (params$annotations) '<script id="annotation-script" src="annotation.js"></script>'`

<!-- close data-nics-theme div -->
</div>
