---
title: "NISRA Technology and Support Lab RAP Skeleton Demo Report"
lang: "en"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 4
    code_download: no
    css: "../style.css"
    includes:
      in_header:
        - "../meta.html"
        - "../consent_head.html"
      after_body:
        - "../cookie_banner.html"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../demo/demo_outputs/",
      output_file = paste0(
        xfun::sans_ext(input),
        "_",
        format(
          Sys.time(),
          "%d-%m-%Y_%H%M"
        ),
        ".html"
      ),
      envir = globalenv()
    )
  })
params:
  annotations: false
---

<div id="cookie-banner"></div>

```{r settings, include=FALSE}
# this chunk sets chunk options for all chunks in this file
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)
```

```{r setup, warnings = FALSE}
library(here)

source(
  here(
    "code",
    "demo",
    "demo_data_prep.R"
  )
)

source(
  here(
    "code",
    "demo",
    "demo_excel_tables.R"
  )
)
```

<div data-nics-theme="`r nics_theme`">
<!-- these <header> and <main> tags are important for accessibility, they
define landmark areas for screenreaders. -->

<header>

<!-- calling the banner code -->

`r f_banner(title = rmarkdown::metadata$title)`

<!-- Header section -->

`r f_header()`

</header>

<!-- Introduction section -->

<!-- open main div -->

<main id="main">

<!-- calling the borderline code -->

`r f_borderline()`

## 1. Key points

This demo report is included with the RAP Skeleton code. It contains:

* code and files to produce all the styling, tables, charts etc contained in
the report
* some good practice advice
* accessibility advice and features

You can use it:

* as worked examples of code you can copy and edit for your data
* to inspire elements of your own reports
* as a guidance document

Throughout the report are:

* features of elements in the report that may be useful to know
* good practice advice
* accessibility requirements and how they have been addressed
* worked examples using population data

### Features

Banner

* The logos used in the banner are set in order. 
* The NISRA logo goes on the left and the department logo on the right. 
* If this is an Accredited Official Statistics report, this logo will be in the
middle. 
* The code automatically puts each logo in the correct place.


Key points

* The figures and years included below are calculated automatically by the code. 
* These will then update automatically when the data changes.
* These key points are arranged in a two-column layout so the chart wraps below
the text when resized.


### Population change

```{r key_point_2_column, echo = TRUE}
point_1 <- paste0(
  "In ",
  earliest_year,
  " about ",
  round_half_up(
    ni_earliest_total,
    2
  ),
  " million people lived in Northern Ireland. By ",
  latest_year,
  ", this had grown to about ",
  round_half_up(
    ni_latest_total,
    0
  ),
  " million people. This means the population increased by ",
  round_half_up(
    ni_earliest_latest_change,
    0
  ),
  " percent."
)

point_2 <- paste0(
  "Children and young people under 25 make up the largest age group ",
  "in Northern Ireland."
)

point_3 <- paste0(
  "Derry City and Strabane saw the biggest population growth ",
  "compared with the other Local Government Districts (LGDs)."
)

change_chart <- here(
  "code",
  "demo",
  "demo_data",
  "images",
  "mini_population.png"
)

div(
  class = "row",
  div(
    class = "col-sm-6",
    style = "margin-top: 10px; font-size: 90%;",
    tags$ul(
      tags$li(point_1),
      tags$li(point_2),
      tags$li(point_3)
    )
  ),
  div(
    class = "col-sm-6",
    img(
      src = change_chart,
      alt = paste0(
        "Column chart showing population growth in Northern Ireland ",
        "between ",
        earliest_year,
        " and ",
        latest_year,
        "."
      )
    )
  )
)
```

`r f_borderline()`

## 2. Introduction

```{r introduction_sub_report, child='02_introduction.Rmd' }
```

`r f_borderline()`

## 3. Infographics and images

```{r infographics_sub_report , child='03_infographics.Rmd'}
```

`r f_borderline()`

## 4. Tables

```{r tables_sub_report , child='04_tables.Rmd'}
```

`r f_borderline()`

## 5. Charts

```{r charts_sub_report  , child='05_charts.Rmd'}
```

`r f_borderline()`

## 6. Maps (h2)

```{r maps_sub_report , child='06_maps.Rmd'}
```

`r f_borderline()`

## 7. Download data {#download-data}

```{r download_data_sub_report , child='07_download_data.Rmd'}
```

`r f_borderline()`

## 8. Notes for readers 

```{r notes_for_readers_sub_report, child='08_notes_for_readers.Rmd'}
```

`r f_borderline()`

## 9. Text presentation

```{r text_presentation_sub_report, child='09_text_presentation.Rmd'}
```

`r f_borderline()`

## 10. Useful accessibility links

```{r useful_links_sub_report, child='10_useful_links.Rmd'}
```

`r f_borderline()`

## 11. Accredited official statistics and official statistics

```{r accredited_stats_sub_report, child='11_accredited_stats.Rmd'}
```

`r f_borderline()`

## 12. Contact Details 


The function called f_contact can also be used to add contact details
automatically. This helps make reports look the same and easier for people to
use.

Email addresses should all be lowercase jo.bloggs@nisra.gov.uk and it is good
practice to include a named individual (some exemptions apply).

`r f_contact()`

### Accessibility contact 

Please contact Dissemination Branch for assistance with accessibility
requirements or alternative formats. Contact details are:

<strong>Email:</strong> [info\@nisra.gov.uk](mailto:info@nisra.gov.uk){.email}

<strong>Telephone:</strong> +44 (0)300 200 7836

Dissemination Branch\
NISRA\
Colby House\
Stranmillis Court\
BELFAST\
BT9 5RR

<!-- close main div -->

</main>

```{r render-footer, results='asis', echo=FALSE}
f_footer()
```


<script>
  // Create array of alt texts for map shapes
  const map_alt = `r if (exists("map_alt")) {map_alt} else {"[]"}`;
</script>


`r if (params$annotations) '<script id="annotation-script" src="annotation.js"></script>'`

```{js}
// Accessibility function to remove map shapes from tab order, add tabs to table
// of contents and add empty alt text to map legend

function tabOrder() {

  // Removes tabs from individual map shapes and appends alt texts

  var shapes = document.getElementsByClassName(
    "leaflet-marker-icon leaflet-zoom-animated leaflet-interactive"
  );

  for (var i = 0; i < shapes.length; i++) {
    shapes[i].tabIndex = -1;
    shapes[i].alt = map_alt[i];
  }

  // Make side menu tabbable

  let headings = document.getElementsByClassName(
    "tocify-item list-group-item"
  );

  let sections = document.getElementsByClassName(
    "section"
  );

  for (let i = 0; i < headings.length; i++) {
    headings[i].innerHTML =
      "<a href='#" +
      sections[i].id +
      "'>" +
      headings[i].innerHTML +
      "</a>";
  }

  // Remove white space at bottom

  const space_observer = new MutationObserver(() => {
    let space = document.getElementsByClassName(
      "tocify-extend-page"
    );

    for (let i = 0; i < space.length; i++) {
      space[i].remove();
    }
  });

  space_observer.observe(
    document.getElementsByClassName("main-container")[0],
    {
      childList: true,
      subtree: true
    }
  );
  
  // Hide entire legend from screen readers

  var legendContainer = document.querySelector(
    ".info.legend"
  );

  if (legendContainer) {
    legendContainer.setAttribute(
      "aria-hidden",
      "true"
    );
  }

  // Add empty alt text to each legend image

  var legendImages = document.querySelectorAll(
    ".info.legend img"
  );

  for (var j = 0; j < legendImages.length; j++) {
    legendImages[j].setAttribute(
      "alt",
      ""
    );

    legendImages[j].setAttribute(
      "aria-hidden",
      "true"
    ); // Extra assurance
  }
}

// Instruct web browser to run above function when loaded
window.onload = tabOrder;

```

<!-- close data-nics-theme div -->
</div>
